### 2018-04-13 Writeup/ms.tex

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Word-Referent Identification Under Multimodal Uncertainty},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Word-Referent Identification Under Multimodal Uncertainty}

  \shorttitle{Word Identification Under Multimodal Uncertainty}

  \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C. Frank\textsuperscript{1}}

  % \def\affdep{{"", ""}}%
  % \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Department of Psychology, Stanford University  }

  \authornote{
    Abdellah Fourtassi

    Department of Psychology

    Stanford University

    50 Serra Mall

    Jordan Hall, Building 420

    Stanford, CA 94301

    Correspondence concerning this article should be addressed to Abdellah
    Fourtassi, Postal address. E-mail:
    \href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}
  }

  \abstract{Identifying a spoken word in a referential context requires both the
ability to process and integrate multimodal input and the ability to
reason under uncertainty. How do these tasks interact with one another?
We introduce a task that allows us to examine how adults identify words
under joint uncertainty in the auditory and visual modalities. We
propose an ideal observer model of the task which provides an account of
how auditory and visual cues are combined optimally. Model predictions
are tested in three experiments where word recognition is made under two
kinds of uncertainty: category ambiguity and/or distorting noise. In all
cases, the optimal model explains much of the variance in human mean
judgments. In particular, when the signal is not distorted with noise,
participants weight the auditory and visual cues optimally, that is,
according to the relative reliability of each modality. But when one
modality has noise added to it, human perceivers systematically prefer
the unperturbed modality to a greater extent than the optimal model
does. The study provides a formal framework which helps us understand
precisely how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergy in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

    \indent Word count: X
  }

  \usepackage[sortcites=false,sorting=none]{biblatex}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

Language uses symbols expressed in one modality (e.g., the auditory
modality, in the case of speech) to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee}. Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead.

How does the listener deal with uncertainty to identify the speaker's
intended word? One rigorous way to approach this question is through
conducting an \emph{ideal observer} analysis. This research strategy
provides a characterization of the task/goal and shows what the optimal
performance should be under this characterization.\footnote{It is, thus,
  a general instance of the rational approach to cognition (Anderson,
  1990). It can also be seens as an instance Marr's computational level
  of analysis.} When there is uncertainty in the input, the ideal
observer performs an optimal probabilistic inference. For example, in
order to recognize an ambiguous linguistic input, the model uses all
available probabilistic knowledge in order to maximize the accuracy of
this recognition. When the task is well specified, the ideal observer
model can be seen as a theoretical upper limit on performance. Thus, it
is supposed to be used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources.

The ideal observer analysis has had a tremendous impact not only on
speech related research (Clayards, Tanenhaus, Aslin, \& Jacobs, 2008;
Feldman, Griffiths, \& Morgan, 2009; Kleinschmidt \& Jaeger, 2015;
Norris \& McQueen, 2008), but also on many other disciplines in the
cognitive sciences (for reviews, see Chater \& Manning, 2006; Knill \&
Pouget, 2004; Tenenbaum, Kemp, Griffiths, \& Goodman, 2011). In
particular, using this research strategy, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (VOT) that differentiated similar words (e.g.,
\enquote{beach} and \enquote{peach}). They found that humans were
sensitive to these probabilistic cues and their judgments closely
reflected the optimal predictions. In another work, Feldman et al.
(2009) studied the perceptual magnet effect, which is a phenomenon that
involves reduced discriminability near prototypical sounds in the native
language (Kuhl, 1991). They showed that this effect can be explained as
the consequence of optimally solving the problem of perception under
uncertainty. Both Clayards et al. (2008) and Feldman et al. (2009)
explored optimal performance under uncertainty in the auditory modality.
There is, however, extensive evidence that information from the visual
modality, such as the speaker's facial features, also influences speech
understanding (see Campbell, 2008 for a review). Bejjanki, Clayards,
Knill, and Aslin (2011) offered a mathematical characterization of how
probabilistic cues from speech and lip movements can be optimally
combined. They showed that human performance during audio-visual
phonemic labeling was consistent (at least at the qualitative level)
with the behavior of the ideal observer.

This previous research, however, did not systematically study speech
understanding when the visual information is obtained, not through the
speaker's facial features, but through the referential context. In fact,
experimental findings showed that information about the identity of the
semantic referent can be integrated with linguistic information to
resolve lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995). To our knowledge, however, no study offered an ideal
observer analysis of word identification in such context, that is, when
the listener has to combine cues from the sound and the referent.
Imagine, for example, that someone is uncertain whether they heard
\enquote{pea} or \enquote{bee}, does this uncertainty make them rely
more on the referent (e.g., the object being pointed at)? Vice versa, if
they are not sure if they saw a bee or a fly, does it make them rely
more on the sound? More importantly, when input in both modalities is
uncertain to varying degrees, do they weight each modality according to
its relative reliability (which is the optimal strategy), or do they
over-rely on a particular modality (which is a sub-optimal strategy)?

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are integrated:

\noindent First, in the case of audio-visual speech, both modalities
offer information about the same underlying speech category. They may
differ only in terms of their informational reliability. In a
referential context, however, the auditory and visual modalities are
additionally different in terms of the roles they play in the
referential process: the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a review).\footnote{There is a debate as to whether
  speech is privileged for infant and adults for the same reasons.
  Whereas some researchers suggest that speech is privileged for both
  infants and adults because of its ability to refer (e.g., Waxman \&
  Markow, 1995), others suggest that speech might \emph{not} have a
  referential status from the start. Rather, speech might be prefered by
  infants only because of a low level auditory ``overshadowing'' (e.g.,
  Sloutsky \& Napolitano, 2003).} Thus, in a referential context, it is
possible that listeners do not treat the auditory and visual modalities
as equivalent sources of information. Instead, there could be a
sub-optimal bias for the auditory modality beyond what is expected from
informational reliability alone.

\noindent Second in the case of audio-visual speech, the auditory and
visual stimuli are expected to be perceptually correlated. The
expectation for this correlation is such that when there is a mismatch
between the auditory and visual input, people still integrate them into
a unified (but illusory) percept (McGurk \& MacDonald, 1976). In the
case of referential language, however, the multimodal association is by
nature \emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance,
there is no logical/perceptual connection between the sound
\enquote{bee} and the corresponding insect. Moreover, variation in the
way the sound \enquote{bee} is pronounced is generally not expected to
correlate perceptually with variation in the shape (or any other visual
property) in the category of bees. In sum, cue combination in the case
of arbitrary audio-visual associations (word-referent) is likely to be
less automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the
auditory and the visual modality to recognize words in a referential
context. In particular, we study how this combination is performed under
various degrees of uncertainty in both the auditory and the visual
modality. We perform a rational analysis of the task. First we propose
an ideal observer model that performs the combination in an optimal
fashion. Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model (which is fit to human responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

We tested the ideal observer model's predictions in three behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership only. In Experiment 2, we intervened by adding background
noise to the auditory modality, and in Experiment 3, we intervened by
adding background noise to the visual modality. In all experiments,
participants were quantitatively near-optimal, though overall response
precision was slightly lower than expected. Moreover, in Experiment 1
where neither of the modalities was perturbed with background noise,
participants weighed auditory and visual cues according to the relative
reliability predicted by the optimal model. In other words, we found no
evidence for a modality bias towards either the auditory or the visual
modality. However, in Experiment 2 and 3, participants over-relied on
one modality when the other modality was perturbed with additional
noise.

\section{Paradigm and Models}\label{paradigm-and-models}

In this section we, first, briefly introduce the multimodal combination
task. Then we explain how behavior in this paradigm can be characterized
optimally with an ideal observer model.

\subsection{The Audio-Visual Word Recognition
Task}\label{the-audio-visual-word-recognition-task}

We introduce a new task that tests word recognition in a referential
context. We use two visual categories (cat and dog) and two auditory
categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For
each participant, an arbitrary pairing is set between the auditory and
the visual categories, leading to two audio-visual word categories
(e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented
with an audio-visual target (the prototype of the target category),
immediately followed by an audio-visual test stimulus
(Figure~\ref{fig:task}). The test stimulus may differ from the target in
both the auditory and the visual components. After these two
presentations, participants press \enquote{same} or \enquote{different.}

\begin{figure}

{\centering \includegraphics[width=400px]{pictures/task}

}

\caption{Overview of the task. In the audio-visual condition, participants are first presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test. The test may differ from the target in both the auditory and the visual components. After these two presentations, participants press `same' (i.e., the same category as the target) or `different' (not the same category). The auditory-only and visual-only conditions are similar to the audio-visual condition, except that only the sounds are heard, or only the pictures are shown, respectively.}\label{fig:task}
\end{figure}

This paradigm is adapted from a previous task (Sloutsky \& Napolitano,
2003), which has been used with both children and adults to probe
audio-visual encoding (see Robinson \& Sloutsky, 2010 for a review). In
the testing phase of the original task, participants are asked whether
or not the two audio-visual presentations are \emph{identical}. In the
current study, we are interested, rather, in the categorization, i.e.,
determining whether or not two similar tokens are members of the same
phonological/semantic category. Therefore, testing in our task is
category-based: Participants are asked to press \enquote{same} if they
think the second item (the test) belongs to the same category as the
first (target) (e.g., dog-/aba/), even if there is a slight difference
in the sound, in the referent, or in both. They are instructed to press
\enquote{different} only if they think that the second stimulus was an
instance of the other category (cat-/ada/). The task also includes
trials where pictures are hidden (audio-only) or where sounds are muted
(visual-only). These unimodal trials provide us with the participants'
evaluation of the probabilistic information present in the auditory and
visual categories. As we shall see, these unimodal distributions are
used as inputs to the optimal cue combination model.

\subsection{Optimal Model}\label{optimal-model}

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---this probabilistic information cannot be characterized with
\emph{sensory noise/variability}, only. Indeed, our task involves
responses over categorical variables (phonemes and concepts), and
therefore, the optimal model in our case should take into account, not
only the noise variability around an individual perceptual estimate, but
also its \emph{within category variability}, i.e., the uncertainty
related to whether this perceptual estimate belongs to a given category
(see also Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki et al., 2011).
In what follows, we describe a model that accounts for both type of
variability. First, we describe the model in the simplified case of
categorical variability only. Second, we augment this simplified model
to account for sensory noise.

\subsubsection{Categorical variability}\label{categorical-variability}

We assume that both the auditory categories (i.e., /aba/ and /ada/) and
the visual categories (cat and dog) are distributed along a single
acoustic and semantic dimension, respectively (Figure~\ref{fig:model}).
Moreover, we assume that all categories are normally distributed.
Formally speaking, if \(A\) denotes an auditory category (/ada/ or
/aba/), then the probability that a point \(a\) along the acoustic
dimension belongs to the category \(A\) is
\[ p(a | A) \sim  N(\mu_A, \sigma^2_A) \] where \(\mu_A\) and
\(\sigma^2_A\) are respectively the mean and the variance of the
auditory category. Similarly, the probability that a point \(v\) along
the visual dimension belongs to the category \(V\) is
\[ p(v | V) \sim  N(\mu_V, \sigma^2_V) \] where \(\mu_V\) and
\(\sigma^2_V\) are the mean and the variance of the visual category. An
audio-visual signal \(w=(a,v)\) can be represented as a point in the
audio-visual space. These audio-visual tokens define bivariate
distributions in the bi-dimentional space. We call these bivariate
distributions \emph{Word categories}, noted \(W\), and are distributed
as follows: \[ p(w | W) \sim  N(M_W, \Sigma_W) \] where
\(M_W=(\mu_A, \mu_V)\) and \(\Sigma_W\) are the mean and the covariance
matrix of the word category. The main assumption of the model is that
the auditory and visual variables are independent (i.e., uncorrelated),
so the covariance matrix is simply: \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

\begin{figure}[!h]
\includegraphics[width=\textwidth]{pictures/model} \caption{Illustration of the model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word type $W_1$ or from the word type $W_2$. Decision threshold is where the guessing probability is 0.5.}\label{fig:model}
\end{figure}

\noindent This assumption simply says that, given a word-object mapping,
e.g., \(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid
assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

Now we turn to the crucial question of modeling how the optimal decision
should proceed given the probabilistic (categorical) information in the
auditory and the visual modalities, as characterized above. We have two
word categories: dog-/aba/ (\(W_1\)) and cat-/ada/
(\(W_2\)).\footnote{This mapping is randomized in the experiments.} When
making decisions, participants can be understood as choosing one of
these two word categories (Figure~\ref{fig:model}). For an ideal
observer, the probability of choosing category 2 when presented with an
audio-visual instance \(w=(a,v)\) is the posterior probability of this
category: \[
p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}
\] Using our assumption that the cues are uncorrelated, we have:
\[p(w | W) = p(a,v| W) = p(a| A)p(v| V)\] Under this assumption, the
posterior probability reduces to the following formula (see Appendix 1
for the details of the derivation):

\begin{equation}
 p(W_2 | w)=\frac{1}{1+(1+b)\exp(\beta_0+\beta_aa+\beta_vv)}
\end{equation}

where \[1+b=\frac{p(W_1)}{p(W_2)}\]
\[\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

\[\beta_a=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\]
\[\beta_v=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}.\]

The parameter \(b\) represents the differential between the categories'
prior probabilities. However, since the identity of word categories is
randomized across participants, \(b\) measures, rather, a response bias
to \enquote{same} if \(b > 0\), and a response bias to
\enquote{different} if \(b < 0\). We expect a general bias towards
answering \enquote{different} because of the categorical nature of our
same-different task: When two items are ambiguous but perceptually
different, participants might have a slight preference for
\enquote{different} over \enquote{same}. As for the means, their values
are fixed, and they correspond to the most typical tokens in our
stimuli. Finally, observations from each modality (\(a\) and \(v\)) are
weighted in Equation 1 according to their reliability:
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}.\]

\subsubsection{Sensory variability}\label{sensory-variability}

So far, we only accounted for categorical variability. For instance, if
the speaker generates a target production \(a_t\) from an auditory
category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the ideal model
assumes that it has direct access to this production token (i.e.,
\(a=a_t\)), and that all uncertainty is about the category membership of
this token. However, we might also want to account for noise in the
brain and/or in the environment. For example, the observer might not
have access to the exact produced target, but only to the target
perturbed by noise. If we assume this noise to be normally distributed,
that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{N_A})\), then integrating
over \(a_t\) leads to the following simple expression:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})\] Similarly, in
the case of sensory noise in the visual modality, we get
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

\[\beta_a \propto \frac{1}{\sigma^2_{A}+\sigma^2_{N_A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V} +\sigma^2_{N_V}}.\]

\subsubsection{Optimal cue combination}\label{optimal-cue-combination}

Equation 1 provides the optimal model's predictions for how
probabilities that characterize uncertainty in the auditory and the
visual modalities can be combined to make categorical decisions.
Parameters' estimates of the probability distributions in each modality
are derived by fitting unimodal posteriors to the participants'
responses in the unimodal conditions, i.e., the condition where only the
sounds are heard or only the pictures are seen
(Figure~\ref{fig:task}).\footnote{Further technical detail about model fitting in the unimodal conditions will be given in the method section of Experiment 1}
Using these derived parameters, the optimal model makes predictions
about responses in the bimodal condition where participants both hear
the sounds and see the pictures.

\subsubsection{Auditory and Visual
baselines}\label{auditory-and-visual-baselines}

The predictions of the optimal model will be compared to two baselines.
The first baseline is a visual model which assumes that participants
rely only on visual information, and an auditory model, which assumes
that participants rely only on auditory information. More precisely,
these baseline models assumes that the participants' responses in the
bimodal condition will not be different from their response in either
the visual-only or the auditory-only condition. If, as we expect, the
participants rely on both the auditory and the visual modalities to make
decision in the bimodal condition, the optimal model would explain more
variance in human responses than the visual or the auditory model do.

\subsection{Descriptive model and analysis of
sub-optimality}\label{descriptive-model-and-analysis-of-sub-optimality}

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. The parameters of this model are
fit to actual responses in the bimodal condition. If the referential
task induces sub-optimality (due, for instance, to the arbitrary nature
of the sound-object association), then the descriptive model should
explain more variance than the optimal model does.

Comparison of the optimal and the descriptive models allows us, not only
to quantify how much people deviate from optimality, but also to
understand precisely the nature of this deviation. Let \(\sigma^2_{A}\)
and \(\sigma^2_{V}\) be the values of the variances used in the optimal
model (derived from the unimodal conditions), and \(\sigma^2_{Ab}\) and
\(\sigma^2_{Vb}\) be the values observed through the descriptive model
in the bimodal condition. Deviation from optimality is measured in two
ways. First, we measure the change in the values of the variance
specific to each modality, that is, how \(\sigma^2_{A}\) compares to
\(\sigma^2_{Ab}\), and how \(\sigma^2_{V}\) compares to
\(\sigma^2_{Vb}\). Second, we measure changes in the proportion of the
visual and auditory variances, i.e., we examine how
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\) compares to
\(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). The first measure allows us to
test if response precision changes for each modality when we move from
the unimodal to the bimodal conditions. The second allows us to test the
extent to which the weighting scheme follows the prediction of the
optimal model. The reason we used the proportion of the variances as a
measure of cross-modal weighting is because this proportion corresponds
to the slope\footnote{Or more precisely the absolute value of the slope}
of the decision threshold in the audio-visual space
(Figure~\ref{fig:model}). The decision threshold is defined as the set
of values in this audio-visual space along which the posterior is equal
to 0.5. Formally speaking, the decision threshold has the following
form:

\[v=-\frac{\sigma^2_V}{\sigma^2_A}a+v_0\]

If the absolute value of the slope derived from the descriptive model is
greater than that of the optimal model, the corresponding shift in the
decision threshold indicates that participants have a preference for the
auditory modality in the bimodal case. Similarly, a smaller absolute
value of the slope would lead to a preference for the visual modality.
The limit cases are when there is exclusive reliance on the auditory cue
(a vertical line), and where there is exclusive reliance on the visual
(a horizontal line).

There are three possible ways human responses can deviate from
optimality. These scenarios are illustrated in Figure~\ref{fig:subOptim}
and are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Both variances may increase, but their proportion remains the same.
  That is, \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), but
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} \approx \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimality would be due to increased randomness in
  human responses in the bimodal condition. However, this randomness
  would not affect the relative weighting of both modalities, i.e.,
  participants would still weigh modalities according to the relative
  reliability predicted by the optimal model.
\item
  The auditory variance increases at a higher rate. That is,
  \(\sigma^2_{Ab} \gg \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimally would consist not only in participants
  being more random in the bimodal condition, but also in having a
  systematic preference for the visual modality, even after accounting
  for informational reliability.
\item
  The visual variance increases at a higher rate. That is,
  \(\sigma^2_{Vb} \gg \sigma^2_{V}\), and
  \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  This case is the reverse of case 2, i.e., in addition to increased
  randomness in the bimodal condition, there is a systematic preference
  for the auditory modality, even after accounting for informational
  reliability.
\end{enumerate}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{pictures/sub-optimal} \caption{Illustration using simulated data showing the example of a prediction made by the optimal model (top), and the three possible ways human participants can deviate from this prediction (bottom). These cases are the following: 1) The variance increases equally for both modalities, but the weighting scheme (characterized by the decision threshold) is optimal, 2) The auditory variance increases at a higher rate, leading to a preference for the auditory modality, and 3) The visual variance increases at a higher rate, leading to a preference for the visual modality.}\label{fig:subOptim}
\end{figure}

We compared these models to human responses in three experiments. In
Experiment 1, we studied the case where bimodal uncertainty was due to
ambiguity in terms of category membership, without any additional
background noise. In Experiment 2 and 3 we added background noise on top
of ambiguity in category membership.

\section{Experiment 1}\label{experiment-1}

In this Experiment, we start with testing the predictions in the case
where uncertainty is due to categorical variability only (i.e.,
ambiguity in terms of category memebership). We do not add any external
noise to the background and we assume that internal sensory noise is
negligible compared to categorical variability
(\(\sigma^2_{A} \gg \sigma^2_{N_A}\), and
\(\sigma^2_{V} \gg \sigma^2_{N_V}\)). Thus, we use the following cue
weighting scheme:

\[\beta_a \propto \frac{1}{\sigma^2_{A} + \sigma^2_{N_A}} \approx  \frac{1}{\sigma^2_{A} }\]
\[\beta_v \propto \frac{1}{\sigma^2_{V} + \sigma^2_{N_V}} \approx  \frac{1}{\sigma^2_{V} }.\]

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. They were paid at an hourly rate
of \$6/hour. Participants were excluded if they reported having
experienced a technical problem of any sort during the online experiment
(N=14), or if they had less than 50\% accurate responses on the
unambiguous training trials (N=6). The final sample consisted of N = 80
participants.\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

\subsubsection{Stimuli}\label{stimuli}

For auditory stimuli, we used the continuum introduced in Vroomen,
Linden, Keetels, Gelder, and Bertelson (2004), a 9-point /aba/--/ada/
speech continuum created by varying the frequency of the second (F2)
formant in equal steps. We selected 5 equally spaced points from the
original continuum by keeping the endpoints (prototypes) 1 and 9, as
well as points 3, 5, and 7 along the continuum. For visual stimuli, we
used a cat/dog morph continuum introduced in Freedman, Riesenhuber,
Poggio, and Miller (2001). From the original 14 points, we selected 5
points as follows: we kept the item that seemed most ambiguous (point
8), the 2 preceding points (i.e., 7 and 6) and the 2 following points
(i.e., 9 and 10). The 6 and 10 points along the morph were quite
distinguishable, and we took them to be our prototypes.

\subsubsection{Design and Procedure}\label{design-and-procedure}

We told participants that an alien was naming two objects: a dog, called
\enquote{aba} in the alien language, and a cat, called \enquote{ada}. In
each trial, we presented the first object (the target) on the left side
of the screen simultaneously with the corresponding sound. For each
participant, the target was always the same (e.g., dog-/aba/). The
second sound-object pair (the test) followed on the other side of the
screen after 500ms and varied in its category membership. For both the
target and the test, visual stimuli were present for the duration of the
sound clip (\(\sim\) 800ms). We instructed participants to press
\enquote{S} for same if they thought the alien was naming another
dog-/aba/, and \enquote{D} for different if they thought the alien was
naming a cat-/ada/. We randomized the sound-object mapping (e.g.,
dog-/aba/, cat-/ada/) as well as the identity of the target (dog or cat)
across participants.

The first part of the experiment trained participants using only the
prototype pictures and the prototype sounds (12 trials, 4 each from the
bimodal, audio-only, and visual-only conditions). After completing
training, we instructed participants on the structure of the task and
encouraged them to base their answers on both the sounds and the
pictures (in the bimodal condition). There were a total of 25 possible
combinations in the bimodal condition, and 5 in each of the unimodal
conditions. Each participant saw each possible trial twice, for a total
of 70 trials/participant. Trials were blocked by condition and blocks
were presented in random order. The experiment lasted around 15 minutes.

\subsubsection{Model fitting details}\label{model-fitting-details}

\paragraph{Unimodal condition}\label{unimodal-condition}

Remember that data in this conditions allow us to derive the variances
of both the auditory and the visual categories, and that these variances
are used to make predictions about bimodal data (in the visual and
auditory baselines as well as in the optimal model). These individual
variances were derived as follows (we explain the derivation for the
auditory-only case, but the same applies for the visual-only case). We
use the same Bayesian reasoning as we did in the derivation of the
bimodal model: When presented with an audio instance \(a\), the
probability of choosing the sound category 2 (that is, to answer
\enquote{different}) is the posterior probability of this category
\(p(A_2|a)\). If we assume that both sound categories have equal
variances, the posterior probability reduces to:

\[p(A_2 | a)=\frac{1}{1+(1+b_A)\exp(\beta_{a0}+\beta_aa)}\]

with \(\beta_a=\frac{\mu_{A_1}-\mu_{A_2}}{\sigma^2_{A}}\) and
\(\beta_{a0}=\frac{\mu^2_{A_2}-\mu^2_{A_1}}{2\sigma^2_{A}}\). \(b_A\) is
the response bias in the auditory-only condition. For this model (as
well as all other models in this study), we fixed the values of the
means to be the end-points of the corresponding continuum, since these
points are the most typical instances in our stimuli. Thus, we have
\(\mu_{A1}=0\) and \(\mu_{A2}=4\) (and similarly \(\mu_{V1}=0\), and
\(\mu_{V2}=4\)). This leaves us with two free parameters: the bias
\(b_A\) and the variance \(\sigma^2_{A}\). To determine the values of
these parameters, we fit the unimodal posterior to human data in the
unimodal case.

\paragraph{Bimodal condition}\label{bimodal-condition}

In this condition, only the descriptive model is fit to the data, using
the expression of the posterior (Equation 1). Since the values of the
means are fixed, we have 3 free parameters: the variances for the visual
and the auditory modalities, respectively, and \(b\), the response bias.
The visual and auditory baselines as well as the optimal model are not
fit to bimodal data, but their predictions are tested against these
bimodal data. All these normative models use the variances derived from
the unimodal data and the bias term derived from the bimodal data.

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant. Instead, models were constructed with data collapsed across
all participants. That being said, the distribution of responses from
individual participants will also be studied. The fit was done with a
nonlinear least squares regression using the NLS package in R (Bates \&
Watts, 1988). We computed the values of the parameters, as well as their
95\% confidence intervals, through non-parametric bootstrap (using 10000
iterations).

\subsection{Results and analysis}\label{results-and-analysis}

\subsubsection{Unimodal conditions}\label{unimodal-conditions}

Average categorization judgments and best fits are shown in
Figure~\ref{fig:unimodal}. The categorization function of the auditory
condition was slightly steeper than that of the visual condition,
meaning that participants perceived the sound tokens slightly more
categorically and whih higher certainty than they did with the visual
tokens. For the auditory modality, we obtained the following
values:\footnote{all CIs in the paper are 95\% confidence intervals.}
\(b_A=\) -0.20 {[}0.02, -0.38{]} and \(\sigma^2_A=\) 2.04 {[}1.66,
2.53{]}. For the visual modality, we obtained \(b_V=\) -0.12 {[}0.06,
-0.28{]} and \(\sigma^2_V=\) 3.33 {[}2.83, 3.92{]}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/unimodal-1} \caption{Human responses in the unimodal conditions. Points represent the proportion of `different' to `same' responses in the auditory-only condition (left), and visual-only condition (right). Error bars are 95\% confidence intervals. Solid lines represent best fits.}\label{fig:unimodal}
\end{figure}

\subsubsection{Bimodal condition}\label{bimodal-condition-1}

\paragraph{Normative models}\label{normative-models}

Figure~\ref{fig:bimodal} compares the predictions of the normative
models against human responses. The visual, auditory and optimal model
explained, respectively, 30\%, 67\%, and 89\% of total variance in mean
responses.

\paragraph{Descriptive model}\label{descriptive-model}

In the descriptive model, all parameters are fit to human responses in
the bimodal condition. We found \(b=\) -0.34 {[}-0.28, -0.39{]},
\(\sigma^2_{Ab}=\) 4.96 {[}4.58, 5.40{]} and \(\sigma^2_{Vb}=\) 7.06
{[}6.40, 7.84{]}. Note that the variance of both the auditory and visual
modalities increased compared to the unimodal conditions.

\noindent The descriptive model explained 95\% of total variance.
However, since the descriptive model was fit to the same data, there is
a risk that this high correlation is due to overfitting. To examine this
possibility, we cross-validated the model using half the responses to
predict the other half (averaging across 1000 random partitions). The
predictive power of the model remained very high (\(r^2\)=0.93).

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/bimodal-1} \caption{Human responses vs. models' predictions in the bimodal condition. Shape represents auditory distance from the target, and color represents visual distance from the target.}\label{fig:bimodal}
\end{figure}

\paragraph{Cue combination and Modality
preference}\label{cue-combination-and-modality-preference}

We next analyzed if cue combination was performed in an optimal way, or
if there was a systematic preference for one modality when making
decisions in the bimodal condition. As explained above, modality
preference can be characterized formally as a deviation from the
decision threshold predicted by the optimal model. Figure~\ref{fig:bias}
(top) shows both the decision threshold derived from the descriptive
model (in black) and the decision threshold predicted by the optimal
model (in red). The deviation from optimality is compared to two
hypothetical cases of modality preference (dotted lines). We found that
the descriptive and optimal decision thresholds were almost identical.
Indeed, non-parametric resampling of the data showed no evidence of a
deviation from the optimal prediction (Figure~\ref{fig:bias}, bottom).

\subsection{Discussion}\label{discussion}

Overall, we found that the optimal model explained much of the variance
in the mean judgments, and largely more than what can be explained with
the auditory or the visual models alone. Moreover, the high value of the
coefficient of determination in the optimal model (\(r^2\)=0.89)
suggests that the population was near-optimal. However, we see in
Figure~\ref{fig:bimodal} that the mean responses deviated systematically
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This is due to the increase in the
value of the variance associated with each modality. Note however that,
despite this increase in randomness, our analysis of modality preference
showed that the relative values of these variances were not different
(Figure~\ref{fig:bias}), meaning that there was no evidence for a
modality preference. Thus, 1) There was a simultaneous increase in the
values of the auditory and visual variances in the bimodal condition
compared to the unimodal condition, meaning that the bimodal input lead
to an increase in response randomness, and 2) this increased randomness
did not affect the relative weighting of both modalities, i.e., the
population was weighting modalities according to the relative
reliability predicted by the optimal model. This situation corresponds
to the first case of sub-optimally described in
Figure~\ref{fig:subOptim}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/individual-1} \caption{Histograms of the values of the visual variance relative to the auditory variance in Experiment 1. Light color represents the values derived from each individual participant, and dark color represents simulated data sampled from the descriptive model.}\label{fig:individual}
\end{figure}

As we noted earlier, the model addresses the question of optimality at
the population level. However, it is important to know how individual
responses are distributed. In fact, one could think of an extreme case
where optimality at the population level would be misleading. Imagine,
for instance, that in the bimodal condition half the participants relied
exclusively on the visual modality, whereas the other half relied
exclusively on the auditory modality. This case could still lead to an
aggregate behaviour which appears optimal, but this optimality would be
spurious.

To examine this possibility, we consider the distribution of individual
cross-modal weighting in the bimodal condition (i.e.,
\(\frac{\sigma^2_{Vb}}{\sigma^2_{Ab}}\)). Using a factor of 10 as a
cut-off, we found that 5 participants relied almost exclusively on the
visual modality, and 12 relied almost exclusively on the auditory
modality. The percentage of both cases was relatively small compared to
the total number of participants (21.25\%). An additional number of
participants (N=7) relied on both modalities, but provided noisy
responses which lead to negative variances (probably due to mistaking
\enquote{same} for \enquote{different} or vice versa). When these
outliers were removed, the distribution had a rather unimodal shape
(Figure~\ref{fig:individual}). This finding indicates that the
population's near optimality is not spurious, but based mostly on
genuine cue combination at the individual level.

As a second analysis, we asked whether the observed variance in the
individual distribution was due to mere sampling errors or whether it
corresponded to a real between-subject variability. We simulated
individual responses from the posterior distribution whose parameters
were fit to the population as a whole (i.e., the descriptive posterior).
The resulting distribution is shown in Figure~\ref{fig:individual}. For
ease of comparison, the simulated distribution was superimposed to the
real distribution. We found that the real distribution was broader
(\(sd=\) 2.24) than the simulated distribution (\(sd=\) 1.17),
indicating that there was a real between-subject variation beyond
sampling errors. This means that the participants varied in terms of how
they weighted modalities: Compared to the predictions of the global
descriptive model, some participants relied more on the auditory
modality, whereas others relied more on the visual modality.

In Experiment 1, we tested word recognition when there was multimodal
uncertainty in terms of category membership only. In real life, however,
tokens can undergo distortions due to noisy factors in the environment
(e.g., car noise in the background, blurry vision in a foggy
weather,..). In Experiment 2 and 3, we explore this additional level of
uncertainty.

\section{Experiment 2}\label{experiment-2}

In this Experiment, we explored the effect of added noise on
performance. We tested a case where the background noise was added to
the auditory modality. We were interested to know if participants would
treat this new source of uncertainty as predicted by the optimal model,
that is, according to the following weighting scheme
\[\beta_a \propto \frac{1}{\sigma^2_{A}+\sigma^2_{N_A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}.\] The alternative hypothesis
is that noise in one modality leads to a systematic preference for the
non-noisy modality.

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

A sample of 100 participants was recruited online through Amazon
Mechanical Turk. We used the same exclusion criteria as in Experiment 1.
7 participants were excluded because they had less than 50\% accurate
responses on the unambiguous training trials. The final sample consisted
of N = 93 participants.

\subsubsection{Stimuli and Procedure}\label{stimuli-and-procedure}

We used the same visual stimuli as in Experiment 1. We also used the
same auditory stimuli, but we convolved each item with Brown noise of
amplitude 1 using the free sound editor Audacity (2.1.2). The average
signal-to-noise ratio was - 4.4 dB. The procedure was exactly the same
as in the previous experiment, except that the test stimuli (but not the
target) were presented with the new noisy auditory stimuli.

\subsection{Results and analysis}\label{results-and-analysis-1}

\subsubsection{Unimodal conditions}\label{unimodal-conditions-1}

We fit a model for each modality. For the auditory modality, our
parameter estimates were \(b_A=\) -0.18 {[}-0.05, -0.30{]} and
\(\sigma^2_A+\sigma^2_N=\) 4.70 {[}4.03, 5.55{]}. For the visual
modality, we found \(b_V=\) -0.24 {[}-0.10, -0.36{]} and \(\sigma^2_V=\)
3.93 {[}3.43, 4.55{]}. Figure~\ref{fig:unimodal} shows responses in the
unimodal conditions as well as the corresponding best fits. The visual
data is a replication of the visual data in Experiment 1. As for the
auditory data, in contrast to Experiment 1, responses were flatter,
showing more uncertainty.

\subsubsection{Bimodal condition}\label{bimodal-condition-2}

\paragraph{Normative models}\label{normative-models-1}

Figure~\ref{fig:bimodal} compares the predictions of the visual,
auditory and optimal models to human responses. These normative models
explained, respectively, 77\%, 21\%, and 91\% of total variance in mean
judgements. Note that, in contrast to Experiment 1, the visual model
explained more variance than the auditory model did.

\paragraph{Descriptive model}\label{descriptive-model-1}

We estimated \(b=\) -0.38 {[}-0.33, -0.42{]},
\(\sigma^2_{Ab}+\sigma^2_{Nb}=\) 9.84 {[}8.75, 11.27{]}, and
\(\sigma^2_{Vb}=\) 5.21 {[}4.84, 5.64{]}. The fit explained 0.97\% of
total variance. Cross-validation using half the responses to predict the
other half yielded \(r^2 =\) 0.95.

\paragraph{Modality preferences}\label{modality-preferences}

Figure~\ref{fig:bias} (top) shows that the participants' decision
threshold deviated from optimality, and that this deviation was biased
towards the visual modality (the non-noisy modality). Indeed
non-parametric resampling of the data showed a decrease in the value of
the slope in the descriptive model compared to the optimal model
(Figure~\ref{fig:bias}, bottom).

\subsection{Discussion}\label{discussion-1}

We found, similar to Experiment 1, that the population was generally
near optimal (\(r^2 =\) 0.91), and that the optimal model explained more
variance than the auditory or the visual models alone. We also found a
similar discrepancy from the optimal model as precision dropped for both
the auditory and the visual modalities. As for the weighting scheme used
by participants, contrary to Experiment 1 where modalities were weighted
according to their relative reliability, we found in this experiment
that the visual modality had a greater weight than what was expected
from its relative reliability. This situation corresponds to the second
case of sub-optimally described in Figure~\ref{fig:subOptim}.

We were also interested in whether noise in the auditory modality lead
more participants to rely exclusively on the visual modality at the
individual level. Using the same cut-off as in Experiment 1 (a factor of
10), the percentage of participants who relied exclusively on either
modalities was 34.41\%, which is much higher than the percentage
obtained in Experiment 1 (21.25\%). Moreover, the subset of participants
relying exclusively on the visual modality (compared to those who relied
exclusively on the auditory modality) increased from 29.41\% in
Experiment 1 to 68.75\% in Experiment 2, indicating that noise in the
auditory modality prompted more participants to rely exclusively and
disproportionately on the visual modality.

In Experiment 2, we tested the case of added background noise to the
auditory modality. In Experiment 3, we test the case of added noise to
the visual modality.

\section{Experiment 3}\label{experiment-3}

In this Experiment, we added background noise to the visual modality.
Similar to Experiment 2, we were interested to know if participants
would treat this new source of uncertainty as predicted by the optimal
model, that is, according to the following weighting scheme:
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}+\sigma^2_{N_V}}.\] The
alternative hypothesis is that, just like noise in the auditory modality
lead to a preference for the visual input in Experiment 2, noise in the
visual modality would lead to a preference for the auditory input.

\subsection{Methods}\label{methods-2}

\subsubsection{Participants}\label{participants-2}

A planned sample of 100 participants was recruited online through Amazon
Mechanical Turk. We used the same exclusion criteria as in both previous
experiments. N=2 participants were excluded because they reported having
a technical problem, and N=10 participants were excluded because they
had less than 50\% accurate responses on the unambiguous training
trials. The final sample consisted of N = 88 participants.

\subsubsection{Stimuli and Procedure}\label{stimuli-and-procedure-1}

We used the same auditory stimuli as in Experiment 1. We also used the
same visual stimuli, but we blurred the tokens using the free image
editor GIMP (2.8.20). We used a Gaussian blur with a
radius\footnote{A features that modulates the intensity of the blur} of
10 pixels. The experimental procedure was exactly the same as in the
previous Experiments.

\subsection{Results and analysis}\label{results-and-analysis-2}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/bias-1} \caption{Modality preference is characterized as a deviation from the optimal decision threshold. A) The decision thresholds of both the optimal and the descriptive models (solid red and black lines respectively). Deviation from optimality is compared to two hypothetical cases of modality preference. In these cases, deviation from  optimality is due to over-lying on the visual or the auditory input (green and blue dotted lines, respectively) by a factor of 2. B) The value of the decision threshold's slope derived from the descriptive model relative to that of the optimal model. Error bars represent 95\% confidence intervals over the distribution obtained through non-parametric resampling.}\label{fig:bias}
\end{figure}

\subsubsection{Unimodal conditions}\label{unimodal-conditions-2}

For the auditory modality, our parameter estimates were \(b_A=\) -0.24
{[}-0.04, -0.42{]} and \(\sigma^2_A=\) 1.94 {[}1.61, 2.33{]}. For the
visual modality, we found \(b_V=\) 0.11 {[}0.27, -0.03{]} and
\(\sigma^2_V+\sigma^2_N=\) 13.00 {[}9.92, 18.94{]}.
Figure~\ref{fig:unimodal} shows responses in the unimodal conditions as
well as the corresponding fits. The auditory data is a replication of
the auditory data in Experiment 1. As for the visual data, we found
that, in contrast to Experiment 1 and 2, responses were flatter, showing
much more uncertainty.

\subsubsection{Bimodal condition}\label{bimodal-condition-3}

\paragraph{Normative models}\label{normative-models-2}

Figure~\ref{fig:bimodal} compares the predictions of the visual,
auditory and optimal models to human responses. These normative models
explained, respectively, 1\%, 98\%, and 97\% of total variance in the
mean judgements.

\paragraph{Descriptive model}\label{descriptive-model-2}

We estimated \(b=\) -0.35 {[}-0.29, -0.40{]}, \(\sigma^2_{Ab}=\) 3.00
{[}2.75, 3.25{]}, and \(\sigma^2_{Vb}+\sigma^2_{Nb}=\) 39.42 {[}25.06,
98.96{]}. The fit explained 97\% of total variance. Cross-validation
using half the responses to predict the other half yielded \(r^2=\)
0.96.

\paragraph{Modality preferences}\label{modality-preferences-1}

Participants' decision threshold suggested a preference for the auditory
modality (the non-noisy modality). Indeed non-parametric resampling of
the data showed an increase in the value of the slope in the descriptive
model compared to the optimal model (Figure~\ref{fig:bias}).

\subsection{Discussion}\label{discussion-2}

We found that the optimal model accounted for almost all the variance
(\(r^2 =\) 0.97). However, whereas in previous experiments the optimal
model explained more variance than the auditory or the visual models,
here the auditory model explained at least as much variance (\(r^2 =\)
0.98). Thus, though participants were still sensitive to variation in
the noisy visual data in the unimodal condition, they tended to ignore
this information in the bimodal condition, and relied almost exclusively
on the non-noisy auditory modality. The reason why we saw this (floor)
effect when we added noise to the visual modality (Experiment 3), and
not when we added noise to the auditory modality (Experiment 2), is the
fact that our visual stimuli were originally perceived less
categorically and with less certainty than the auditory stimuli. This
made it more likely for the visual categorization function to become
flat and uninformative after a few drops in precision due to noise on
the one had, and to the additional randomness induced by the bimodal
presentation on the other hand.

The general finding corresponds to the third case of sub-optimality
described in Figure~\ref{fig:subOptim}. Indeed, precision dropped for
both modalities in the bimodal condition compared to the unimodal
condition. But the drop was much greater for the visual modality,
resulting in a much lower weight assigned to it than what is expected
from its reliability. Therefore, just like participants over-relied on
the visual modality when the auditory modality was noisy (Experiment 2),
they also over-relied on the auditory modality when the visual modality
was noisy (Experiment 3).

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exclusive}The percentage of participants who relied exclusively on either the visual modality or the auditory modality, using a factor of 10 as a cut-off (e.g., we consider that a participant relied exclusively on the visual modality when their auditory variance is a at least 10 times larger than their visual variance). We show the percentage compared to the total number of participants in each Experiment (`Total'). From this subset of participants, we show the percentage of those who relied on the  auditory modality (`Auditory'), and the percentage of those who relied on the visual  modality (Visual').}
\begin{tabular}{llll}
\toprule
Experiment & \multicolumn{1}{c}{Total} & \multicolumn{1}{c}{Auditory} & \multicolumn{1}{c}{Visual}\\
\midrule
Exp1 & 21.25 & 70.59 & 29.41\\
Exp2 & 34.41 & 31.25 & 68.75\\
Exp3 & 38.64 & 94.12 & 5.88\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

The percentage of participants who relied exclusively on either the
visual modality or the auditory modality was 38.64\%, which is closer to
the percentage of Experiment 2, except that now almost all of them
relied on the auditory modality (94.12\%). For ease of comparison, Table
\ref{tab:exclusive} provides a summary of the numbers across the three
Experiments.

\section{General Discussion}\label{general-discussion}

When identifying a spoken word under uncertainty, one often needs to
make the most of the available cues. Some previous work studied optimal
behavior under uncertainty from the auditory input only (e.g., Clayards
et al., 2008; Feldman et al., 2009), and others studied optimality under
multimodal uncertainty in auditory speech and visual facial features
(e.g. Bejjanki et al., 2011). The current work explored, for the first
time, the case of word identification under uncertainty in speech (word
form) and the visual word \emph{referent}. More specifically, we
conducted an ideal observer analysis of the task whereby a model
provided predictions about how information from each modality should be
combined in an optimal fashion. The predictions of the model were tested
in a series of three experiments where instances of both the form and
the meaning were ambiguous with respect to their category membership
only (Experiment 1), when instances of the form were perturbed with
additional background noise (Experiment 2), and when instances of the
referent were perturbed with additional visual noise (Experiment 3).

In all Experiments, we found many patterns of optimal behaviour.
Quantitatively speaking, the optimal model accounted, respectively, for
89\%, 91\%, and 97\% of the variance in mean responses. When compared to
the predictions of the visual or the auditory models, participants
generally relied on both modalities to make their decisions in the
bimodal condition. Indeed, in Experiment 1 and 2, the optimal model
accounted for more variance in mean responses than the auditory or the
visual models did. In Experiment 3, participants appeared to rely on one
modality, but this was likely a floor effect, due to the fact that noise
made the visual input barely perceivable. In Experiment 1, which did not
involve background noise, participants not only relied on both
modalities, but generally weighted these modalities according to the
prediction of the optimal model, that is, according to their relative
reliability. At the individual level, however, we found evidence of a
between-subject variation: Some participants relied slightly more on the
visual modality, whereas others relied slightly more on the auditory
modality.

We documented two major cases of sub-optimality. First, in all
Experiments, the variance associated with each modality increased in the
bimodal condition compared to the unimodal conditions. This means that
participants responded slightly more randomly in the bimodal condition
than they did in the unimodal conditions. This finding contrasts with
research on multisensory integration where associations tend to lead to
a higher precision (e.g., Ernst \& Banks, 2002). Nevertheless, there is
a crucial difference between these two situations (besides the obvious
difference in terms of the models used). Research on multisensory
integration (of which audio-visual speech is arguably an instance) deals
with redundant multimodal cues, and these cues are integrated into a
unified percept. In contrast, the word-referent association is usually
arbitrary and, in particular, the cues are not expected to be correlated
perceptually. Therefore the observer cannot form a unified percept,
rather, it must encode information separately from both modalities and
retain this encoding through the decision making process. Retaining two
separate cues at the same time instead of forming one unified percept
(as in multisensory integration of redundant cues), or instead of
retaining only one cue (as in the unimodal case), is likely to place
extra-demand on cognitive resources, which, in turn, can cause general
performance to drop. Indeed, there is evidence that cognitive load has a
detrimental effect on word recognition. This phenomenon can be due to a
reduction in perceptual acuity (e.g., Mattys \& Wiget, 2011).

Some previous research found a similar case of suboptimal behavior. For
instance, studies that explored the identification of ambiguous, newly
learned pairs of word-referent associations all reported what appears to
be a decrease in speech perception acuity in both children (Stager \&
Werker, 1997) and adults (Pajak, Creel, \& Levy, 2016). Recently, Hofer
and Levy (2017) provided a probabilistic model of this phenomenon. In
agreement with the finding in the current study, Hofer and Levy (2017)
characterized the apparent reduction in perceptual acuity as an increase
in the noise variance of the auditory modality. Our finding, besides
providing more evidence to this documented fact, suggests that the
reduction in perceptual acuity may occurs simultaneously in both the
auditory \emph{and} the visual modalities.

The second case of sub-optimality is related to how participants weighed
the cues from the visual and the auditory modalities in a noisy context.
In contrast to Experiment 1 where the combination was indistinguishable
from the optimal prediction, results of Experiment 2 and 3 which both
involved background noise in one modality, showed that participants had
a systematic preference for the other (non-noisy) modality. From
previous empirical studies, we know that when the speech signal is
degraded, people tend to compensate by relying more on other sources of
information such as the accompanying visual cues (i.e., lip movements)
or the semantic/syntactic context (see Mattys, Davis, Bradlow, \& Scott,
2012 for a review). However, and generally speaking, these studies do
not differentiate between an optimal compensatory strategy (i.e.,
relying more on the alternative source while using all information still
available in the distorted signal), and a sub-optimal strategy (i.e.,
relying more on the alternative source while ignoring at least some of
the information still available in the distorted signal). The formal
approach followed in this paper allowed us to tease apart these two
possibilities, and the analysis supports the sub-optimal compensatory
strategy: The preference for the non-noisy modality is above and beyond
what can be explained by the relative reliability alone, meaning that
the participants tend to ignore at least part of the information still
available in the noisy modality.

This second case of sub-optimal behavior is possibly related to the fact
that language understanding under degraded conditions is cognitively
more taxing than language understanding under normal conditions (e.g.,
Ronnberg, Rudner, Lunner, \& Zekveld, 2010). This fact can lead to a
bias against the more noisy cue. One could also explain this phenomenon
in terms of the metacognitive experience about the fluency with which
information is processed. The perceived perceptual fluency (e.g., the
ease with which a stimulus' physical identity can be identified) can
affect a wide variety of human judgements (see Schwarz, 2004 for a
review). In particular, variables that improve fluency tends to increase
liking/preference (Reber, Winkielman, \& Schwarz, 1998). In our case,
the subjective experience of lower fluency in the noisy modality might
cause people to underestimate information that can be extracted from
this modality, especially when presented simultaneously with a higher
fluency alternative.

An important question to ask is how the combination mechanism---as
revealed in our controlled study---scales up to real life situations.
Note that in order to test audio-visual cue combination under
uncertainty, we had to use a case of double ambiguity, that is, a case
where both the word forms (\enquote{ada}-\enquote{aba}) and the
referents (cat-dog) were similar and, thus, confusable. However, to what
extent does such case occur in real languages? Cross-linguistic corpus
analyses suggest that lexical encoding tends, surprisingly, towards
double ambiguity in many languages (Dautriche, Mahowald, Gibson, \&
Piantadosi, 2017; Monaghan, Shillcock, Christiansen, \& Kirby, 2014;
Tamariz, 2008). For instance, Dautriche et al. (2017) analyzed 100
languages and found that words that are similar phonologically tend to
be similar semantically as well, beyond what could be explained by
chance. These studies suggest that the case of double uncertainty,
though perhaps not pervasive, could be a real issue in language as it
increase the probability of confusability for many
words.\linebreak Besides the case of double ambiguity intrinsic to
language, there are two situations where our mechanism might play a
significant role. The first is when ambiguity in both the form and/or
the referent is induced by an external noisy context even when these
forms and referents are not confusable in normal situations. The second
case is that of early word learning, and we will discuss this case in
more detail in what follows.

Though we only tested adults in this paper, the problem of word
recognition under uncertainty, as well as the need to make the most of
ambiguous cues, is a particularly pressing issue for children. In fact,
whereas adults are mostly faced with uncertainty in the \emph{input},
children have to deal with the additional uncertainty that results from
their early unrefined \emph{representations} of both phonological and
semantic categories. For example, upon hearing a noisy instance of
\enquote{bee}, adults may have to decide whether the speaker intended to
say \enquote{pea} or \enquote{bee}, but children can additionally be
uncertain whether \enquote{bee} is a different word from \enquote{pee}
(as opposed to, say, a valid within category variation), especially if
these similar sounding words are newly learned (Creel, 2012; Merriman \&
Schuster, 1991; Stager \& Werker, 1997; Swingley, 2016; White \& Morgan,
2008). Though new word form representation can be shown to be
differentiated under some circumstances (e.g., Yoshida, Fennell,
Swingley, \& Werker, 2009), this differentiation is still not mature
enough and is probably noisier than the adult-like representation and/or
encoded with lower confidence (see Swingley, 2007).

At the semantic level, early representations have, similarly, an
intrinsically fragile and uncertain status. For example, upon seeing a
bee in a foggy weather, adults may be uncertain if they saw a bee or a
fly. But on top of this perceptual uncertainty, children may not be
certain if the semantic category being named is that of bees and only
bees, or if it includes other small flying insects like flies and
beetles. In fact, though children can be fast at learning a first
approximation of a given word's referent (Carey \& Bartlett, 1978), the
refinement of this early approximation into a mature semantic category
is a slow and gradual process (see also Bion, Borovsky, \& Fernald,
2013; Carey, 2010; Fernald, Perfors, \& Marchman, 2006; McMurray, Horst,
\& Samuelson, 2012). Among other things, children have to enrich this
early representation with new features, and revise its extension in the
light of new referential exposures.

Thus, uncertainty in the representation associated with one modality
(e.g., a bee and a fly) can be mitigated through the possibly more
differentiated representations associated with the other modality (e.g.,
the sound \enquote{bee} is acoustically different from the sound
\enquote{fly}). That being said, a multi-modal cue combination strategy
might help children not only recognize an individual word instance, but
also refine the underlying phonological and semantic representations in
the process. Previous research in early word learning has---whether
implicitly or explicitly---largely treated the process of learning form
and of learning meaning as independent. However, the developmental data
reviewed above shows that children do not wait to have completed the
acquisition of form to start learning meanings, and that both form and
meaning representations develop, rather, in a parallel fashion. A few
studies pointed to the possibility of an interaction between sound and
meaning in early acquisition. For instance, Waxman and Markow (1995)
showed that labeling various objects with the same name helps infants
form the broad semantic category (but see Sloutsky \& Napolitano, 2003).
Vice versa, Yeung and Werker (2009) showed that pairing similar sounds
with different objects help infants pay attention to subtle phonological
contrasts. The present study proposes a first step towards a formal
framework where accounts of sound-meaning synergy in development can be
unified and further explored. For example, one could imagine that,
initially, visual and auditory categories have relatively large noise
variances, and that development consists in reducing the values of the
variances through a mutually constraining process as further multimodal
data accumulates.

One limitation of the work is that we used simplified stimuli. For the
auditory modality, we used speech categories that varied along a single
acoustic dimension. While this dimension might be sufficient to
recognize words in our specific case, in general the speech signal may
be more complex, varying along several acoustic/phonetic dimensions.
Additionally, these dimensions may be highly variable due to various
kinds of speaker and context differences. The same thing can be said
about the referential stimuli. Here we used a continuum along a single
morph dimension in order to construct a multimodal input where the
auditory and visual components have symmetrical properties. Though such
morph is not the exact visual variability that people would encounter in
they daily lives, it allowed us to precisely test the role of auditory
and visual information in the cue combination process. Parameterizing
semantic dimensions is a notoriously difficult problem, but morphs have
been used in previous research as a reasonable proxy (Freedman et al.,
2001; Havy \& Waxman, 2016; Sloutsky \& Fisher, 2004). It is an open
question as to whether people use the same strategy in controlled
laboratory conditions, as in more naturalistic settings where they have
to deal with various levels of variability. However an answer to this
question is likely to involve a multifaceted approach,
involving---besides laboratory experiments---analyses of corpora with a
more realistic multimodal input (e.g., Fourtassi, Schatz, Varadarajan,
\& Dupoux, 2014; Harwath, Torralba, \& Glass, 2016; B. C. Roy, Frank,
DeCamp, \& Roy, 2015).

In conclusion, this work studied the mechanism of word identification
under uncertainty in both the word form and the word referent. To our
knowledge, this is the first study that performs an ideal observer
analysis of this task. We found people to be near optimal in their cue
combination: They weighted each modality according to its relative
reliability. However, they also showed patterns of sub-optimality
especially when the stimuli were perturbed with additional background
noise. Though the present study did not directly address the issue of
early word learning, it provides a framework where developmental
questions can also be investigated. For instance, future work should
explore whether children, like adults, use probabilistic cues from both
the auditory and the visual input to recognize ambiguous words, the
extent to which they combine these cues in an optimal fashion, and
whether these combination help them with refining their early
phonological and semantic representations.

\section{Appendix 1: derivation of the posterior (Equation
1)}\label{appendix-1-derivation-of-the-posterior-equation-1}

For an ideal observer, the probability of choosing category 2 when
presented with an audio-visual instance \(w = (a, v)\) is the posterior
probability of this category:

\[p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}\]

Which reduces to:

\[p(W_2 | w)=\frac{1}{1+\frac{p(w|W_1)}{p(w|W_2)} \frac{p(W_1)}{p(W_2)}}\]
In order to further simplify the quantity \(\frac{p(w|W_1)}{p(w|W_2)}\),
we use our assumption that the cues are uncorrelated:
\[p(w | W) = p(a,v| W) = p(a| A)p(v| V)\] Using the \(\log\)
transformation, we get:

\[ \ln(\frac{p(w |W_1)}{p(w|W_2)})=\ln(\frac{p(a|W_1)}{p(a|W_2)})+\ln(\frac{p(v|W_1)}{p(v|W_2)}) \]
Under the assumption that the categories are normally distributed and
that, within each modality, the categories have equal variances, we get
(after simplification):

\[\ln(\frac{p(a|W_1)}{p(a|W_2)})=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\times a+ \frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}\]

and similarly:

\[\ln(\frac{p(v|W_1)}{p(v|W_2)})=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}\times v+ \frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

When putting all these terms together, we obtain this final expression
for the posterior:
\[p(W_2 | w)=\frac{1}{1+(1+b)\exp(\beta_0+\beta_aa+\beta_vv)}\]

where

\[1+b=\frac{p(W_1)}{p(W_2)}\]
\[\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

\[\beta_a=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\]
\[\beta_v=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}.\]

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-anderson90}{}
Anderson, J. R. (1990). \emph{The adaptive character of thought}.
Hillsdale, NJ: Erlbaum.

\hypertarget{ref-Bankieris17}{}
Bankieris, K. R., Bejjanki, V., \& Aslin, R. N. (2017). Sensory
cue-combination in the context of newly learned categories.
\emph{Scientific Reports}, \emph{7}(1), 10890.

\hypertarget{ref-bates88}{}
Bates, D., \& Watts, D. (1988). \emph{Nonlinear regression analysis and
its applications}. Wiley.

\hypertarget{ref-bejjanki2011}{}
Bejjanki, V., Clayards, M., Knill, D., \& Aslin, R. (2011). Cue
integration in categorical tasks: Insights from audio-visual speech
perception. \emph{PLoS ONE}, \emph{6}.

\hypertarget{ref-bion2013}{}
Bion, R. A., Borovsky, A., \& Fernald, A. (2013). Fast mapping, slow
learning: Disambiguation of novel word-object mappings in relation to
vocabulary learning at 18, 24, and 30 months. \emph{Cognition},
\emph{126}(1), 39--53.

\hypertarget{ref-Campbell2008}{}
Campbell, R. (2008). The processing of audio-visual speech: Empirical
and neural bases. \emph{Philosophical Transactions of the Royal Society
of London B: Biological Sciences}, \emph{363}(1493), 1001--1010.

\hypertarget{ref-carey2010}{}
Carey, S. (2010). Beyond fast mapping. \emph{Language Learning and
Development}, \emph{6}(3), 184--205.

\hypertarget{ref-carey1978b}{}
Carey, S., \& Bartlett, E. (1978). Acquiring a single new word. In
\emph{Proceedings of the Stanford Child Language Conference} (Vol. 15,
pp. 17--29).

\hypertarget{ref-chater06}{}
Chater, N., \& Manning, C. D. (2006). Probabilistic models of language
processing and acquisition. \emph{Trends in Cognitive Sciences},
\emph{10}, 335--344.

\hypertarget{ref-clayard08}{}
Clayards, M., Tanenhaus, M., Aslin, R., \& Jacobs, R. (2008). Perception
of speech reflects optimal use of probabilistic speech cues.
\emph{Cognition}, \emph{108}.

\hypertarget{ref-Creel2012}{}
Creel, S. (2012). Phonological similarity and mutual exclusivity:
On-line recognition of atypical pronunciations in 3--5-year-olds.
\emph{Developmental Science}, \emph{15}(5), 697--713.

\hypertarget{ref-dautriche17}{}
Dautriche, I., Mahowald, K., Gibson, E., \& Piantadosi, S. (2017).
Wordform similarity increases with semantic similarity: An analysis of
100 languages. \emph{Cognitive Science}, \emph{41}(8), 2149--2169.

\hypertarget{ref-Eberhard1995}{}
Eberhard, K., Spivey-Knowlton, M. J., Sedivy, J. C., \& Tanenhaus, M.
(1995). Eye movements as a window into real-time spoken language
comprehension in natural contexts. \emph{Journal of Psycholinguistic
Research}, \emph{24}(6), 409--436.

\hypertarget{ref-ernst02}{}
Ernst, M. O., \& Banks, M. S. (2002). Humans integrate visual and haptic
information in a statistically optimal fashion. \emph{Nature},
\emph{415}(6870), 429--433.

\hypertarget{ref-feldman2009}{}
Feldman, N., Griffiths, T., \& Morgan, J. (2009). The influence of
categories on perception: Explaining the perceptual magnet effect as
optimal statistical inference. \emph{Psychological Review},
\emph{116}(4), 752--782.

\hypertarget{ref-fernald2006}{}
Fernald, A., Perfors, A., \& Marchman, V. (2006). Picking up speed in
understanding: Speech processing efficiency and vocabulary growth across
the 2nd year. \emph{Developmental Psychology}, \emph{42}(1), 98--116.

\hypertarget{ref-fourtassi2014b}{}
Fourtassi, A., Schatz, T., Varadarajan, B., \& Dupoux, E. (2014).
Exploring the relative role of bottom-up and top-down information in
phoneme learning. In \emph{Proceedings of the 52nd annual meeting of the
association for computational linguistics (volume 2: Short papers)}
(Vol. 2, pp. 1--6).

\hypertarget{ref-freedman2001}{}
Freedman, D., Riesenhuber, M., Poggio, T., \& Miller, E. and. (2001).
Categorical representation of visual stimuli in the primate prefrontal
cortex. \emph{Science}, \emph{291}.

\hypertarget{ref-Geisler2003}{}
Geisler, W. S. (2003). Ideal observer analysis. In \emph{The visual
neurosciences} (pp. 825--837)). Cambridge, MA: MIT Press.

\hypertarget{ref-greenberg1957}{}
Greenberg, J. (1957). \emph{Essays in linguistics}. Chicago: University
of Chicago Press.

\hypertarget{ref-harwath2016}{}
Harwath, D., Torralba, A., \& Glass, J. (2016). Unsupervised learning of
spoken language with visual context. In \emph{Advances in neural
information processing systems} (pp. 1858--1866).

\hypertarget{ref-havy2016}{}
Havy, M., \& Waxman, S. R. (2016). Naming influences 9-month-olds'
identification of discrete categories along a perceptual continuum.
\emph{Cognition}, \emph{156}, 41--51.

\hypertarget{ref-hillenbrand1995}{}
Hillenbrand, J., Getty, L. A., Clark, M. J., \& Wheeler, K. (1995).
Acoustic characteristics of american english vowels. \emph{Journal of
the Acoustical Society of America}, \emph{97}.

\hypertarget{ref-hofer2017}{}
Hofer, M., \& Levy, R. (2017). Modeling Sources of Uncertainty in Spoken
Word Learning. In \emph{Proceedings of the 39th Annual Meeting of the
Cognitive Science Society}.

\hypertarget{ref-kleinschmidt2015}{}
Kleinschmidt, D. F., \& Jaeger, T. F. (2015). Robust speech perception:
Recognize the familiar, generalize to the similar, and adapt to the
novel. \emph{Psychological Review}, \emph{148}.

\hypertarget{ref-Knill04}{}
Knill, D., \& Pouget, A. (2004). The bayesian brain: The role of
uncertainty in neural coding and computation. \emph{Trends in
Neurosciences}, \emph{27}(12), 712--719.

\hypertarget{ref-kuhl1991}{}
Kuhl, P. K. (1991). Human adults and human infants show a ``perceptual
magnet effect'' for the prototypes of speech categories, monkeys do not.
\emph{Perception \& Psychophysics}, \emph{50}(2), 93--107.

\hypertarget{ref-mattys11}{}
Mattys, S. L., \& Wiget, L. (2011). Effects of cognitive load on speech
recognition. \emph{Journal of Memory and Language}, \emph{65}(2),
145--160.

\hypertarget{ref-mattys12}{}
Mattys, S. L., Davis, M. H., Bradlow, A. R., \& Scott, S. K. (2012).
Speech recognition in adverse conditions: A review. \emph{Language and
Cognitive Processes}, \emph{27}(7-8), 953--978.

\hypertarget{ref-mcgurk1976}{}
McGurk, H., \& MacDonald, J. (1976). Hearing lips and seeing voices.
\emph{Nature}, \emph{264}, 746--748.

\hypertarget{ref-mcmurray2012}{}
McMurray, B., Horst, J. S., \& Samuelson, L. K. (2012). Word learning
emerges from the interaction of online referent selection and slow
associative learning. \emph{Psychological Review}, \emph{119}.

\hypertarget{ref-Merriman91}{}
Merriman, W., \& Schuster, J. (1991). Young children's disambiguation of
object name reference. \emph{Child Development}, \emph{62}(6),
1288--1301.

\hypertarget{ref-Monaghan2014}{}
Monaghan, P., Shillcock, R. C., Christiansen, M. H., \& Kirby, S.
(2014). How arbitrary is language? \emph{Philosophical Transactions of
the Royal Society of London B: Biological Sciences}, \emph{369}(1651).

\hypertarget{ref-Norris08}{}
Norris, D., \& McQueen, J. M. (2008). Shortlist B: A bayesian model of
continuous speech recognition. \emph{Psychological Review},
\emph{115}(2), 357--395.

\hypertarget{ref-pajak2016}{}
Pajak, B., Creel, S., \& Levy, R. (2016). Difficulty in learning
similar-sounding words: A developmental stage or a general property of
learning? \emph{Journal of Experimental Psychology: Learning, Memory,
and Cognition}, \emph{42}(9).

\hypertarget{ref-rahnev2018}{}
Rahnev, D., \& Denison, R. N. (2018). Suboptimality in perceptual
decision making. \emph{Behavioral and Brain Sciences}, 1?107.

\hypertarget{ref-reber98}{}
Reber, R., Winkielman, P., \& Schwarz, N. (1998). Effects of perceptual
fluency on affective judgments. \emph{Psychological Science},
\emph{9}(1), 45--48.

\hypertarget{ref-robinson2010}{}
Robinson, C. W., \& Sloutsky, V. (2010). Development of cross-modal
processing. \emph{Wiley Interdisciplinary Reviews: Cognitive Science},
\emph{1}.

\hypertarget{ref-Ronnberg10}{}
Ronnberg, J., Rudner, M., Lunner, T., \& Zekveld, A. (2010). When
cognition kicks in: Working memory and speech understanding in noise.
\emph{Noise and Health}, \emph{12}(49), 263--269.

\hypertarget{ref-roy2015}{}
Roy, B. C., Frank, M. C., DeCamp, M., P., \& Roy, D. (2015). Predicting
the birth of a spoken word. \emph{Proceedings of the National Academy of
Sciences}, \emph{112}.

\hypertarget{ref-saussure1916}{}
Saussure, F. (1916). \emph{Course in general linguistics.} New York:
McGraw-Hill.

\hypertarget{ref-schwarz2004}{}
Schwarz, N. (2004). Metacognitive experiences in consumer judgment and
decision making. \emph{Journal of Consumer Psychology}, \emph{14}(4),
332--348.

\hypertarget{ref-sloutsky2004}{}
Sloutsky, V., \& Fisher, A. V. (2004). Induction and categorization in
young children: A similarity-based model. \emph{Journal of Experimental
Psychology: General}, \emph{133}(2), 166.

\hypertarget{ref-sloutsky2003}{}
Sloutsky, V., \& Napolitano, A. (2003). Is a picture worth a thousand
words? Preference for auditory modality in young children. \emph{Child
Development}, \emph{74}.

\hypertarget{ref-spivey2002}{}
Spivey, M. J., Tanenhaus, M., Eberhard, K., \& Sedivy, J. C. (2002). Eye
movements and spoken language comprehension: Effects of visual context
on syntactic ambiguity resolution. \emph{Cognitive Psychology},
\emph{45}(4), 447--481.

\hypertarget{ref-stager1997}{}
Stager, C. L., \& Werker, J. F. (1997). Infants listen for more phonetic
detail in speech perception than in word-learning tasks. \emph{Nature},
\emph{388}(6640).

\hypertarget{ref-Swingley2007}{}
Swingley, D. (2007). Lexical exposure and word-form encoding in
1.5-year-olds. \emph{Developmental Psychology}, \emph{43}(2), 454--464.

\hypertarget{ref-Swingley2016}{}
Swingley, D. (2016). Two-year-olds interpret novel phonological
neighbors as familiar words. \emph{Developmental Psychology},
\emph{52}(7), 1011--1023.

\hypertarget{ref-Tamariz2008}{}
Tamariz, M. (2008). Exploring systematicity between phonological and
context-cooccurrence representations of the mental lexicon. \emph{The
Mental Lexicon}, \emph{3}(2).

\hypertarget{ref-Tanenhaus1995}{}
Tanenhaus, M., Spivey-Knowlton, M., Eberhard, K., \& Sedivy, J. (1995).
Integration of visual and linguistic information in spoken language
comprehension. \emph{Science}, \emph{268}(5217), 1632--1634.

\hypertarget{ref-tenenbaum11}{}
Tenenbaum, J., Kemp, C., Griffiths, T., \& Goodman, N. (2011). How to
grow a mind: Statistics, structure, and abstraction. \emph{Science},
\emph{331}(11 March 2011), 1279--1285.

\hypertarget{ref-vouloumanos2014}{}
Vouloumanos, A., \& Waxman, S. (2014). Listen up! Speech is for thinking
during infancy. \emph{Trends in Cognitive Sciences}, \emph{18}(12),
642--646.

\hypertarget{ref-vroomen2004}{}
Vroomen, J., Linden, S. van, Keetels, M., Gelder, B. de, \& Bertelson,
P. (2004). Selective adaptation and recalibration of auditory speech by
lipread information: Dissipation. \emph{Speech Communication},
\emph{44}.

\hypertarget{ref-waxman1995}{}
Waxman, S., \& Markow, D. (1995). Words as invitations to form
categories: Evidence from 12-to 13-month-old infants. \emph{Cognitive
Psychology}, \emph{29}(3), 257--302.

\hypertarget{ref-white2008b}{}
White, K., \& Morgan, J. (2008). Sub-segmental detail in early lexical
representations. \emph{Journal of Memory and Language}, \emph{59}.

\hypertarget{ref-yeung09}{}
Yeung, H., \& Werker, J. (2009). Learning words' sounds before learning
how words sound: 9-month-olds use distinct objects as cues to categorize
speech information. \emph{Cognition}, \emph{113}, 234--243.

\hypertarget{ref-yoshida2009}{}
Yoshida, K., Fennell, C., Swingley, D., \& Werker, J. (2009).
14-month-olds learn similar-sounding words. \emph{Developmental
Science}, \emph{12}.

\end{document}