### 2019-11-14 Writeup/ms_r2.tex

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={How Optimal is Word Recognition Under Multimodal Uncertainty?},
            pdfauthor={Abdellah Fourtassi~\& Michael C. Frank},
            pdfkeywords={Language understanding; audio-visual processing; word learning; speech
perception; computational modeling.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{How Optimal is Word Recognition Under Multimodal Uncertainty?}
    \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C.
Frank\textsuperscript{1}}
    \date{}

\shorttitle{Word Identification Under Multimodal Uncertainty}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\keywords{Language understanding; audio-visual processing; word learning; speech perception; computational modeling.}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\usepackage{xcolor}
\usepackage[sortcites=false,sorting=none]{biblatex}

\authornote{ Abdellah Fourtassi

Department of Psychology

Stanford University

50 Serra Mall

Jordan Hall, Building 420

Stanford, CA 94301

Correspondence concerning this article should be addressed to Abdellah
Fourtassi, Postal address. E-mail:
\href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}}

\abstract{
Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We study how
adults identify novel words under joint uncertainty in the auditory and
visual modalities and we propose an ideal observer model of how cues in
these modalities are combined optimally. Model predictions are tested in
four experiments where recognition is made under various sources of
uncertainty. We found that participants use both auditory and visual
cues to recognize novel words. When the signal is not distorted with
environmental noise, participants weight the auditory and visual cues
optimally, that is, according to the relative reliability of each
modality. In contrast, when one modality has noise added to it, human
perceivers systematically prefer the unperturbed modality to a greater
extent than the optimal model does. This work extends the literature on
perceptual cue combination to the case of word recognition in a
referential context. In addition, this context offers a link to the
study of multimodal information in word meaning learning.

}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\begin{verbatim}
## Warning: package 'bindrcpp' was built under R version 3.4.4
\end{verbatim}

\section{Introduction}\label{introduction}

Language uses symbols expressed in one modality --- the auditory
modality, in the case of speech --- to communicate about the world,
which we perceive through many different sensory modalities. Consider
hearing someone yell \enquote{bee!} at a picnic, as a honey bee buzzes
around the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provides convergent evidence.

However, word identification takes place in a noisy world, and the cues
received through each modality may not provide a definitive answer. On
the auditory side, individual acoustic word tokens are almost always
ambiguous with respect to the particular sequence of phonemes they
represent, which is due to the inherent variability of how a phonetic
category is realized acoustically (Hillenbrand, Getty, Clark, \&
Wheeler, 1995). Moreover, some tokens may be distorted additionally by
mispronunciation or ambient noise. Perhaps the speaker was yelling
\enquote{pea} and not \enquote{bee.} Similarly, a sensory impression may
not be enough to make a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be accomplished through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

As a simplified case study of early word learning, the task of matching
sounds to corresponding visual objects has been studied extensively in
the developmental literature. For example, many studies focus on how
children might succeed in this type of task despite referential
ambiguity (Medina, Snedeker, Trueswell, \& Gleitman, 2011; Pinker, 1989;
L. B. Smith \& Yu, 2008; Suanda, Mugwanya, \& Namy, 2014; Vlach \&
Johnson, 2013; Vouloumanos, 2008; Yurovsky \& Frank, 2015). However,
even when they have learned the exact meaning of a word, observers (both
children and adults) often still find it challenging to recognize which
word the speaker has uttered, especially under noise (Mattys, Davis,
Bradlow, \& Scott, 2012; Peelle, 2018). The purpose of the current study
is thus to explore word recognition by adults under multimodal
uncertainty, focusing on the special case where people have access to
multimodal cues from the auditory speech and the visual referent. In the
General Discussion, we return to the question of how these findings
relate to questions about word learning.

One rigorous way to approach this question is through conducting an
\emph{ideal observer} analysis. This research strategy provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization.\footnote{It is, thus, a general
  instance of the rational approach to cognition (Anderson, 1990),
  instantiating Marr's computational level of analysis (Marr, 1982).}
When there is uncertainty in the input, the ideal observer performs an
optimal probabilistic inference. For example, in order to recognize an
ambiguous linguistic input, the model uses all available probabilistic
knowledge in order to maximize the accuracy of this recognition. The
ideal observer model can be seen as a theoretical upper limit on
performance. It is not so much a realistic model of human performance,
as much as a baseline against which human performance can be compared
(Geisler, 2003; Rahnev \& Denison, 2018). When there is a deviation from
the ideal, it can reveal extra constraints on human cognition, such as
limitations on the working memory or on attentional resources. This
approach has had a tremendous impact not only on speech-related research
(Clayards, Tanenhaus, Aslin, \& Jacobs, 2008; Feldman, Griffiths, \&
Morgan, 2009; Kleinschmidt \& Jaeger, 2015; Norris \& McQueen, 2008) but
also on many other disciplines in the cognitive sciences (for reviews,
see Chater \& Manning, 2006; Knill \& Pouget, 2004; Tenenbaum, Kemp,
Griffiths, \& Goodman, 2011).

Some prior ideal observer studies are closely related to the question we
are addressing in the current work. For instance, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (Voice Onset Time) that differentiated similar
words (e.g., \enquote{beach} and \enquote{peach}). They found that
humans were sensitive to these probabilistic cues and their judgments
closely reflected the optimal predictions. Moreover, Feldman et al.
(2009) studied the perceptual magnet effect, a phenomenon that involves
reduced discriminability near prototypical sounds in the native language
(Kuhl, 1991), showing that this effect can be explained as the
consequence of optimally solving the problem of perception under
uncertainty (see also Kronrod, Coppess, \& Feldman, 2016).

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the predictions of an ideal observer. This
previous research did not, however, study speech understanding when
visual information was obtained through the referential context rather
than through observation of the speaker's face. Although some
experimental findings show that information about the identity of a
referent can be integrated with linguistic information to resolve
lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995), to our knowledge no study has offered an ideal observer
analysis of this task (as we do here).

Combining information between words and visual referents might seem
similar to audio-visual speech integration (e.g., Bejjanki et al.,
2011), but there are at least two fundamental differences between these
two cases, and both can influence the way the auditory and visual cues
are combined.

First, in the case of audio-visual speech, both modalities offer
information about the same underlying speech category. They differ only
in terms of their informational reliability. In a referential context,
however, the auditory and visual modalities play different roles in the
referential process --- the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning} (and these
differences are in addition to possible differences in informational
reliability). Speech is claimed to have a privileged status compared to
other sensory stimuli (Edmiston \& Lupyan, 2015; Lupyan \&
Thompson-Schill, 2012; Vouloumanos \& Waxman, 2014; Waxman \& Gelman,
2009; Waxman \& Markow, 1995), and this privilege is suggested to be
specifically related to the ability to refer (Waxman \& Gelman,
2009).\footnote{There is, however, a debate as to whether speech is
  privileged for children and adults for similar reasons. Whereas some
  researchers suggest that speech is privileged for both children and
  adults because of its ability to refer (e.g., Waxman \& Gelman, 2009),
  others suggest that speech might \emph{not} have a referential status
  from the start. Rather, speech might be preferred by children only
  because of a low level auditory ``overshadowing'' (e.g., Sloutsky \&
  Napolitano, 2003). We do not take a stance on this issue here.} Thus,
in a referential context, it is possible that listeners do not treat the
auditory and visual modalities as equivalent sources of information.
Instead, there could be a (potentially sub-optimal) bias for the
auditory modality beyond what is expected from informational reliability
alone.

Second, in the case of audio-visual speech, the auditory and visual
stimuli are expected to be perceptually correlated. The expectation for
this correlation is strong enough that when there is a mismatch between
the auditory and visual input, they are still integrated into a unified
(but illusory) percept (e.g., the McGurk Effect; McGurk \& MacDonald,
1976). In the case of referential language, however, the multimodal
association is by nature \emph{arbitrary} (Greenberg, 1957; Saussure,
1916). For instance, there is no logical or perceptual connection
between the sound \enquote{bee} and the corresponding insect. Moreover,
variation in the way the sound \enquote{bee} is pronounced is generally
not expected to correlate perceptually with variation in the shape (or
any other visual property) in the category of bees. In sum, cue
combination in the case of arbitrary audio-visual associations
(word-referent) is likely to be less automatic, more effortful, and
therefore less conducive to optimal integration than it is in the case
of perceptually correlated associations (as in audio-visual speech
perception).

\subsection{The current study}\label{the-current-study}

We investigate how cues from the auditory and the visual modality are
combined to recognize novel words in a referential context. In
particular, we study how this combination is performed under various
degrees of uncertainty in both the auditory and the visual modality.
Imagine, for example, that someone is uncertain whether they heard
\enquote{pea} or \enquote{bee.} Does this uncertainty make them rely
more on the referent (e.g., the object being pointed at)? Or, if they
are not sure if they saw a bee or a fly, does this uncertainty make them
rely more on the sound? More importantly, when input in both modalities
is uncertain to varying degrees, do they weight each modality according
to its relative reliability (the optimal strategy), or do they over-rely
on a particular modality?

We begin by proposing an ideal observer model that performs the
combination in an optimal fashion. We then compare the predictions of
the optimal model to human responses. Humans can deviate from the ideal
for several reasons. For instance, as mentioned above, a sub-optimality
can be induced by the privileged status of a particular modality or by
the arbitrariness of the referential association. In order to study
possible patterns of sub-optimality, we compare the optimal normative
model to a descriptive model (which is fit to actual responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

We tested the ideal observer model's predictions in four behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership (in addition to sensory noise). In Experiment 2, we
intervened by adding environmental noise that degraded information from
the auditory modality and in Experiment 3 we intervened by adding
environmental noise that degraded information from the visual modality.
Finally, in Experiment 4, we replicated Experiment 1 with a higher power
design, allowing us to test cue combination at the individual level.

\section{Paradigm and Models}\label{paradigm-and-models}

In this section, we first briefly introduce the multimodal combination
task. Then we explain how behavior in this paradigm can be characterized
optimally with an ideal observer model.

\subsection{The Audio-Visual Word Recognition
Task}\label{the-audio-visual-word-recognition-task}

We introduce an experimental paradigm adapted from a task used by
Sloutsky and Napolitano (2003). The original was used with both children
and adults to probe audio-visual encoding (see Robinson \& Sloutsky,
2010 for review). Here we use a slightly different version to test word
recognition in a referential context. We use two visual categories (cat
and dog) and two auditory categories (/b/ and /d/ embedded in the
minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is
set between the auditory and the visual categories, leading to two
audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each
trial, participants are presented with an audio-visual target (the
prototype of the target category), immediately followed by an
audio-visual test stimulus (Figure~\ref{fig:task}). The test stimulus
may differ from the target in both the auditory and the visual
components. After these two presentations, participants press
\enquote{same} or \enquote{different.}

\begin{figure}

{\centering \includegraphics[width=400px]{pictures/task}

}

\caption{Overview of the task. In the audio-visual condition, participants are first presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test. The test may differ from the target in both the auditory and the visual components. After these two presentations, participants press `same' (i.e., the same category as the target) or `different' (not the same category). The auditory-only and visual-only conditions are similar to the audio-visual condition, except that only the sounds are heard, or only the pictures are shown, respectively.}\label{fig:task}
\end{figure}

In the testing phase of the original task (Sloutsky \& Napolitano,
2003), participants were asked whether or not the two audio-visual
presentations are \emph{identical}. In the current study, we are
interested, rather, in the categorization, i.e., determining whether or
not two similar tokens are members of the same phonological/semantic
category. Therefore, testing in our task is category-based: Participants
are asked to press \enquote{same} if they think the second item (the
test) belongs to the same category as the first (target) (e.g.,
cat-/aba/), even if there is a slight difference in the sound, in the
referent, or in both. They are instructed to press \enquote{different}
only if they think that the second stimulus was an instance of the other
category (cat-/ada/). The task also includes trials where pictures are
hidden (audio-only) or where sounds are muted (visual-only). These
unimodal trials provide us with the participants' evaluation of the
probabilistic information present in the auditory and visual categories.
As we shall see, these unimodal distributions are used as inputs to the
optimal cue combination model.

\subsection{Optimal Model}\label{optimal-model}

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002), which typically studies continuous stimuli (e.g., size,
location), the probabilistic information in our case cannot be
characterized with \emph{sensory noise} only. Since our task involves
responses over categorical variables (phonemes and concepts), the
optimal model should take into account not only the noise variability
around an individual perceptual estimate but also its \emph{categorical
variability}, i.e., the uncertainty related to whether this perceptual
estimate belongs to a given category (see also Bankieris, Bejjanki, \&
Aslin, 2017; Bejjanki et al., 2011). In what follows, we describe a
probabilistic model that accounts for both types of variability. First,
we describe the model in the simplified case of categorical variability
only. Second, we augment this simplified model to account for sensory
and environmental noise.

\subsubsection{Categorical variability}\label{categorical-variability}

We assume that both the auditory categories (i.e., /aba/ and /ada/) and
the visual categories (cat and dog) are distributed along a single
acoustic and semantic dimension, respectively (Figure~\ref{fig:model}).
Moreover, we assume that all categories are normally distributed.
Formally speaking, if \(A\) denotes an auditory category (/ada/ or
/aba/), then the probability that a point \(a\) along the acoustic
dimension belongs to the category \(A\) is
\[ p(a | A) \sim  N(\mu_A, \sigma^2_A) \] where \(\mu_A\) and
\(\sigma^2_A\) are respectively the mean and the variance of the
auditory category. Similarly, the probability that a point \(v\) along
the visual dimension belongs to the category \(V\) is
\[ p(v | V) \sim  N(\mu_V, \sigma^2_V) \] where \(\mu_V\) and
\(\sigma^2_V\) are the mean and the variance of the visual category. An
audio-visual signal \(w=(a,v)\) can be represented as a point in the
audio-visual space. These audio-visual tokens define bivariate
distributions in the bi-dimensional space. We call these bivariate
distributions \emph{Word categories}, noted \(W\), and are distributed
as follows: \[ p(w | W) \sim  N(M_W, \Sigma_W) \] where
\(M_W=(\mu_A, \mu_V)\) and \(\Sigma_W\) are the mean and the covariance
matrix of the word category. The main assumption of the model is that
the auditory and visual variables are independent (i.e., uncorrelated),
so the covariance matrix is simply: \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

\begin{figure}[!h]
\includegraphics[width=\textwidth]{pictures/model} \caption{Illustration of the model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word type $W_1$ or from the word type $W_2$. Decision threshold is where the guessing probability is 0.5.}\label{fig:model}
\end{figure}

\noindent This assumption says that, given a word-object mapping, e.g.,
\(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid assumption in the context of our
task.\footnote{Note that this assumption is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

Now we turn to the crucial question of modeling how the optimal decision
should proceed given the probabilistic (categorical) information in the
auditory and the visual modalities, as characterized above. We have two
word categories: dog-/aba/ (\(W_1\)) and cat-/ada/
(\(W_2\)).\footnote{This mapping is randomized in the experiments.} When
making decisions, participants can be understood as choosing one of
these two word categories (Figure~\ref{fig:model}). For an ideal
observer, the probability of choosing category 2 when presented with an
audio-visual instance \(w=(a,v)\) is the posterior probability of this
category: \[
p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}
\] Using our assumption that the cues are uncorrelated, we have:
\[p(w | W) = p(a,v| W) = p(a| A)p(v| V)\] Under this assumption, the
posterior probability reduces to the following formula (see Appendix 1
for the details of the derivation):

\begin{equation}
 p(W_2 | w)=\frac{1}{1+(1+b)\exp(\beta_0+\beta_aa+\beta_vv)}
\end{equation}

where \[1+b=\frac{p(W_1)}{p(W_2)}\]
\[\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

\[\beta_a=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\]
\[\beta_v=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}.\]

The parameter \(b\) represents the differential between the categories'
prior probabilities. However, since the identity of word categories is
randomized across participants, \(b\) measures, rather, a response bias
to \enquote{same} if \(b > 0\), and a response bias to
\enquote{different} if \(b < 0\). We expect a general bias towards
answering \enquote{different} because of the categorical nature of our
same-different task: When two items are ambiguous but perceptually
different, participants might have a slight preference for
\enquote{different} over \enquote{same.} As for the means, their values
are fixed, and they correspond to the most typical tokens in our
stimuli. Finally, observations from each modality (\(a\) and \(v\)) are
weighted in Equation 1 according to their reliability (that is,
according to the \emph{inverse} of their variance):
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}.\]

\subsubsection{Sensory variability}\label{sensory-variability}

So far, we have only accounted for categorical variability, i.e.,
\(\sigma^2_{A} = \sigma^2_{A_C}\). For instance, if the speaker
generates a target production \(a_t\) from an auditory category
\(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A_C})\), the ideal model assumes
that it has direct access to this production token (i.e., \(a=a_t\)),
and that all uncertainty is about the category membership of this token.
However, we might also want to account for internal noise in the brain
and/or external noise in the environment. For example, the observer
might not have access to the exact produced target, but only to the
target perturbed by noise. If we assume this noise to be normally
distributed, that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{A_N})\), then
integrating over \(a_t\) leads to this new expression of the probability
distribution:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A_C}+\sigma^2_{A_N})\] Similarly,
in the case of sensory noise in the visual modality, we get:
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{V_N})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

\[\beta_a \propto \frac{1}{\sigma^2_{A_C}+\sigma^2_{A_N}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V_C} +\sigma^2_{V_N}}.\]

\subsubsection{Optimal cue combination}\label{optimal-cue-combination}

Equation 1 provides the optimal model's predictions for how
probabilities that characterize uncertainty in the auditory and the
visual modalities can be combined to make categorical decisions.
Parameter estimates of the probability distributions in each modality
are derived by fitting unimodal posteriors to the participants'
responses in the unimodal conditions, i.e., the condition where only the
sounds are heard or only the pictures are seen
(Figure~\ref{fig:task}).\footnote{Further technical detail about model fitting in the unimodal conditions will be given in the method section of Experiment 1.}
Using these derived parameters, the optimal model makes predictions
about responses in the bimodal (i.e., audio-visual) condition where
participants both hear the sounds and see the pictures.

\subsubsection{Auditory and Visual
baselines}\label{auditory-and-visual-baselines}

The predictions of the optimal model will be compared to two baselines.
The first baseline is a visual model which assumes that participants
rely only on visual information, and an auditory model, which assumes
that participants rely only on auditory information. More precisely,
these baseline models assume that the participants' responses in the
bimodal condition will not be different from their response in either
the visual-only or the auditory-only condition. However, if the
participants rely on both the auditory and the visual modalities to make
decisions in the bimodal condition, the optimal model would explain more
variance in human responses than the visual or the auditory model does.

\subsection{Descriptive model and analysis of
(sub-)optimality}\label{descriptive-model-and-analysis-of-sub-optimality}

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. It is formally identical to the
normative optimal model (Equation 1), except that the parameters are fit
to actual responses in the bimodal condition. If the referential task
induces sub-optimality (due, for instance, to the arbitrary nature of
the sound-object association), then the descriptive model should explain
more variance than the optimal model does.

Comparison of the optimal and the descriptive models allows us, not only
to quantify how much people deviate from optimality but also to
understand precisely the nature of this deviation. Let \(\sigma^2_{A}\)
and \(\sigma^2_{V}\) be the values of the variances used in the optimal
model (derived from the unimodal conditions), and \(\sigma^2_{Ab}\) and
\(\sigma^2_{Vb}\) be the values observed through the descriptive model
in the bimodal condition. Deviation from optimality is measured in two
ways. First, we measure the change in the values of the variance
specific to each modality, that is, how \(\sigma^2_{A}\) compares to
\(\sigma^2_{Ab}\), and how \(\sigma^2_{V}\) compares to
\(\sigma^2_{Vb}\). Second, we measure changes in the proportion of the
visual and auditory variances, i.e., we examine how
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\) compares to
\(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). The first measure allows us to
test if response precision changes for each modality when we move from
the unimodal to the bimodal conditions. The second allows us to test the
extent to which the weighting scheme follows the prediction of the
optimal model. The reason we used the proportion of the variances as a
measure of cross-modal weighting is that this proportion corresponds to
the slope\footnote{Or more precisely the absolute value of the slope.}
of the decision threshold in the audio-visual space
(Figure~\ref{fig:model}). The decision threshold is defined as the set
of values in this audio-visual space along which the posterior is equal
to 0.5. Formally speaking, the decision threshold has the following
form:

\[v=-\frac{\sigma^2_V}{\sigma^2_A}a+v_0\]

If the absolute value of the slope derived from the descriptive model is
greater than that of the optimal model, the corresponding shift in the
decision threshold indicates that participants have a preference for the
auditory modality in the bimodal case. Similarly, a smaller absolute
value of the slope would lead to a preference for the visual modality.
The limit cases are when there is exclusive reliance on the auditory cue
(a vertical line), and where there is exclusive reliance on the visual
(a horizontal line).

There are three possible ways human responses can deviate from
optimality. These scenarios are illustrated in
Figure~\ref{fig:subOptim}, and are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Both variances may increase, but their proportion remains the same.
  That is, \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), but
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} \approx \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimality would be due to increased randomness in
  human responses in the bimodal condition. However, this randomness
  would not affect the relative weighting of both modalities, i.e.,
  participants would still weigh modalities according to the relative
  reliability predicted by the optimal model.
\item
  The auditory variance increases at a higher rate. That is,
  \(\sigma^2_{Ab} \gg \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimality would consist not only in participants
  being more random in the bimodal condition, but also in having a
  systematic preference for the visual modality, even after accounting
  for informational reliability.
\item
  The visual variance increases at a higher rate. That is,
  \(\sigma^2_{Vb} \gg \sigma^2_{V}\), and
  \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  This case is the reverse of case 2, i.e., in addition to increased
  randomness in the bimodal condition, there is a systematic preference
  for the auditory modality, even after accounting for informational
  reliability.
\end{enumerate}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{pictures/sub-optimal} \caption{Illustration using simulated data showing the example of a prediction made by the optimal model (top), and the three possible ways human participants can deviate from this prediction (bottom). These cases are the following: 1) The variance increases equally for both modalities, but the weighting scheme (characterized by the decision threshold) is optimal, 2) The auditory variance increases at a higher rate, leading to a preference for the auditory modality, and 3) The visual variance increases at a higher rate, leading to a preference for the visual modality.}\label{fig:subOptim}
\end{figure}

We compared these models to human responses in four experiments. In
Experiment 1, we studied the case where bimodal uncertainty was due to
categorical variability and sensory noise. In Experiment 2 and 3 we
added environmental noise to the auditory and the visual modalities,
respectively. Finally, in Experiment 4, we replicated Experiment 1 with
a higher power design in order to test cue combination at the individual
level

\section{Experiment 1}\label{experiment-1}

In this Experiment, we test the predictions of the model in the case
where uncertainty is due to categorical variability (i.e., ambiguity in
terms of category membership) and inherent sensory noise. We do not add
any external noise to the background. Thus, we test the following
(normative) cue weighting scheme.\footnote{In all our analyses, we study
  the combined effect of both categorical and sensory uncertainties in
  the process of cue combination. We do not examine the specific role of
  each source of uncertainty separately. For example, in the auditory
  modality, the quantity \(\sigma^2_{A_C} + \sigma^2_{A_N}\) will be
  treated as a single variable. This variable will be measured in the
  unimodal condition and tested in the bimodal condition.}

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{A_C} + \sigma^2_{A_N}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{V_C} + \sigma^2_{V_N}}.\]

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. Participants were excluded if
they reported having experienced a technical problem of any sort during
the online experiment (\(N =\) 14), or if they had less than 50\%
accurate responses on the unambiguous training trials (\(N =\) 6). The
final sample consisted of \(N =\) 80 participants. All participants
provided informed consent before taking the experiment.
\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

\subsubsection{Stimuli}\label{stimuli}

For auditory stimuli, we used the continuum introduced in Vroomen,
Linden, Keetels, Gelder, and Bertelson (2004), a 9-point /aba/--/ada/
speech continuum created by varying the frequency of the second (F2)
formant in equal steps. We selected 5 equally spaced points from the
original continuum by keeping the endpoints (prototypes) 1 and 9, as
well as points 3, 5, and 7 along the continuum. For visual stimuli, we
used a cat/dog morph continuum introduced in Freedman, Riesenhuber,
Poggio, and Miller (2001). From the original 14 points, we selected 5
points as follows: we kept the item that seemed most ambiguous (point
8), the 2 preceding points (i.e., 7 and 6) and the 2 following points
(i.e., 9 and 10). The 6 and 10 points along the morph were quite
distinguishable, and we took them to be our prototypes.

\subsubsection{Design and Procedure}\label{design-and-procedure}

We told participants that an alien was naming two objects: a dog, called
\enquote{aba} in the alien language, and a cat, called \enquote{ada}. In
each trial, we presented the first object (the target) on the left side
of the screen simultaneously with the corresponding sound. For each
participant, the target was always the same (e.g., dog-/aba/). The
second sound-object pair (the test) followed on the other side of the
screen after 500ms and varied in its category membership. For both the
target and the test, visual stimuli were present for the duration of the
sound clip (\(\sim\) 800ms). We instructed participants to press
\enquote{S} for same if they thought the alien was naming another
dog-/aba/, and \enquote{D} for different if they thought the alien was
naming a cat-/ada/. We randomized the sound-object mapping (e.g.,
dog-/aba/, cat-/ada/) as well as the identity of the target (dog or cat)
across participants.

The first part of the experiment trained participants using only the
prototype pictures and the prototype sounds (12 trials, 4 each from the
bimodal, audio-only, and visual-only conditions). After completing
training, we instructed participants on the structure of the task and
encouraged them to base their answers on both the sounds and the
pictures (in the bimodal condition). There were a total of 25 possible
combinations in the bimodal condition, and 5 in each of the unimodal
conditions. Each participant saw each possible trial twice, for a total
of 70 trials/participant. Trials were blocked by condition and blocks
were presented in random order. The experiment lasted around 15
minutes.\footnote{The experiment can be accessed and played from the
  github repository: \url{https://github.com/afourtassi/WordRec/}}

\subsubsection{Model fitting details}\label{model-fitting-details}

\paragraph{Unimodal conditions}\label{unimodal-conditions}

Remember that data in these conditions allows us to derive the variances
of both the auditory and the visual categories and that these variances
are used to make predictions about bimodal data (in the visual and
auditory baselines as well as in the optimal model). These individual
variances were derived as follows (we explain the derivation for the
auditory-only case, but the same applies to the visual-only case). We
use the same Bayesian reasoning as we did in the derivation of the
bimodal model: When presented with an audio instance \(a\), the
probability of choosing the sound category 2 (that is, to answer
\enquote{different}) is the posterior probability of this category
\(p(A_2|a)\). If we assume that both sound categories have equal
variances, the posterior probability reduces to:

\[p(A_2 | a)=\frac{1}{1+(1+b_A)\exp(\beta_{a0}+\beta_aa)}\]

with \(\beta_a=\frac{\mu_{A_1}-\mu_{A_2}}{\sigma^2_{A}}\) and
\(\beta_{a0}=\frac{\mu^2_{A_2}-\mu^2_{A_1}}{2\sigma^2_{A}}\). \(b_A\) is
the response bias in the auditory-only condition. For this model (as
well as all other models in this study), we fixed the values of the
means to be the end-points of the corresponding continuum, since these
points are the most typical instances in our stimuli. Thus, we have
\(\mu_{A1}=0\) and \(\mu_{A2}=4\) (and similarly \(\mu_{V1}=0\), and
\(\mu_{V2}=4\)). This leaves us with two free parameters: the bias
\(b_A\) and the variance \(\sigma^2_{A}\). To determine the values of
these parameters, we fit the unimodal posterior to human data in the
unimodal case.

\paragraph{Bimodal condition}\label{bimodal-condition}

In this condition, only the descriptive model is fit to the data, using
the expression of the posterior (Equation 1). Since the values of the
means are fixed, we have 3 free parameters: the variances for the visual
and the auditory modalities, respectively, and \(b\), the response bias.
The visual and auditory baselines, as well as the optimal model, are not
fit to the bimodal data, but their predictions are tested against these
bimodal data. All these normative models use the variances derived from
the unimodal data and the bias term derived from the fit to bimodal
data.

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant (but see Experiment 4). Instead, models were constructed
with data collapsed across all participants. The fit was done with a
nonlinear least squares regression using the NLS package in R (Bates \&
Watts, 1988). We computed the values of the parameters using
non-parametric bootstrap (with 10000 iterations).

\begin{table}

\caption{\label{tab:param}Estimates of the models' parameters.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Auditory} & \multicolumn{2}{c}{Visual} & \multicolumn{3}{c}{Bimodal} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5} \cmidrule(l{2pt}r{2pt}){6-8}
Experiment & b\textsubscript{A} & Var\textsubscript{A} & b\textsubscript{V} & Var\textsubscript{V} & b\textsubscript{b} & Var\textsubscript{Ab} & Var\textsubscript{Vb}\\
\midrule
Experiment1 & -0.20 & 2.04 & -0.12 & 3.33 & -0.34 & 4.96 & 7.06\\
Experiment2 & -0.18 & 4.70 & -0.24 & 3.93 & -0.38 & 9.84 & 5.21\\
Experiment3 & -0.24 & 1.94 & 0.11 & 13.00 & -0.35 & 3.00 & 39.42\\
Experiment4 & -0.40 & 1.92 & -0.22 & 3.24 & -0.42 & 4.17 & 7.28\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Results and analysis}\label{results-and-analysis}

\subsubsection{Unimodal conditions}\label{unimodal-conditions-1}

Average categorization judgments and best fits are shown in Figure
\ref{fig:unimodal}. The categorization function of the auditory
condition was slightly steeper than that of the visual condition,
meaning that participants perceived the sound tokens slightly more
categorically and with higher certainty than they did with the visual
tokens. The unimodal models' estimates in each modality are shown in
Table \ref{tab:param}.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_r2_files/figure-latex/unimodal-1} \caption{Human responses in the unimodal conditions across the four experiments. Points represent the proportion of `different' to `same' responses in the auditory-only condition (left), and visual-only condition (right). Error bars are 95\% confidence intervals. Solid lines represent best unimodal posterior fits.}\label{fig:unimodal}
\end{figure}

\subsubsection{Bimodal condition}\label{bimodal-condition-1}

Figure \ref{fig:bimodal} compares the predictions of the normative and
descriptive models against human responses. Remember that the normative
models use the parameters estimated from the unimodal conditions (where
people see input from only one modality) to predict behavior in the
bimodal condition (where people see input from both modalities). The
descriptive model has a similar structure than the optimal model but is
directly fit to human responses in the bimodal condition in order to
allow us to assess deviation from optimality.

We found, by comparing the correlation values, that the optimal model
explained more variance than the visual and auditory models did.
However, the optimal model was not perfect: It explained less variance
than the descriptive model did, which indicates a deviation from
optimality. To investigate this deviation, we compare the parameter
values of the optimal model to the values obtained in the descriptive
model (Table
\ref{tab:param}).\footnote{Note that the descriptive model explained almost all the variance in mean responses, which makes it a reasonable proxy for human real performance in the bimodal condition.}
We note an increase in \emph{both} the auditory and visual variances.
This increase in noise is compatible with the fact that human responses
appear to be pulled towards chance (i.e., the value 0.5) when compared
to the optimal model (see Figure \ref{fig:bimodal}). Below we
investigate if this deviation from optimality can be related to the cue
combination strategy.

\subsubsection{Cue combination}\label{cue-combination}

We analyzed if the cue combination was performed in an optimal way, or
if there was a systematic preference for one modality when making
decisions in the bimodal condition. As explained in Figure
\ref{fig:subOptim}, modality preference can be characterized formally as
a deviation from the decision threshold predicted by the optimal model.
The results in Figure \ref{fig:bias} (top) show both the decision
threshold derived from the descriptive model (in black) and the decision
threshold predicted by the optimal model (in red). We found that the
descriptive and optimal decision thresholds were almost identical.
Indeed, non-parametric resampling of the data showed no evidence of a
deviation from the optimal prediction (Figure~\ref{fig:bias}, bottom).

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_r2_files/figure-latex/bimodal-1} \caption{Human responses vs. Models' predictions in the bimodal condition across the four experiments. Each point represents data from a particular audio-visual matching (corresponding to an instance from the set of 5x5 possible matchings in the audio-visual space). Shape represents auditory distance from the target, and color represents visual distance from the target. Thus, each point is characterized by both  shape and color.}\label{fig:bimodal}
\end{figure}

\subsection{Discussion}\label{discussion}

This experiment studied the way participants combine multimodal
information to recognize novel words. We found that the optimal model
explained more variance than the auditory or the visual models did,
indicating that participants take into account both the auditory and
visual cues when making a decision. That said, Figure \ref{fig:bimodal}
shows that the participants deviated slightly --- but systematically---
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This fact was captured by the
increase in the value of the variance associated with each modality (as
can be noted from Table \ref{tab:param}). Note, however, that despite
this increase in response randomness, our analysis of modality
preference showed that the \emph{relative} values of these variances
were not different (Figure~\ref{fig:bias}), meaning that there was no
evidence for a modality preference.

To sum up, 1) the participants used both the auditory and visual
information, 2) they responded slightly more randomly that what was
predicted, but 3) this increased randomness was general and did not
influence the cue combination strategy, i.e., the participants still
weighted modalities according to their relative reliability as predicted
by the optimal model. This situation corresponds to the first case of
sub-optimality described in Figure~\ref{fig:subOptim}.

In Experiment 1, we tested word recognition when there was multimodal
uncertainty in terms of category membership and perceptual noise. In
real life, however, both sound and visual tokens can undergo distortions
due to noisy factors in the environment (e.g., car noise in the
background, blurry vision in foggy weather). In Experiment 2 and 3, we
explore this additional level of uncertainty.

\section{Experiment 2}\label{experiment-2}

In this Experiment, we explored the effect of added envirnonmental noise
\(\sigma^2_{E}\) on performance. We tested a case where the background
noise was added to the auditory modality. We were interested to know if
participants would treat this new source of uncertainty as predicted by
the optimal model, that is, according to the following weighting scheme:

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{A_C}+\sigma^2_{A_N} + \sigma^2_{A_E}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{V_C}+\sigma^2_{V_N}}.\]

The alternative hypothesis is that noise in one modality leads to a
systematic preference for the non-noisy modality.

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

A sample of \(N =\) 100 participants was recruited online through Amazon
Mechanical Turk. We used the same exclusion criteria as in Experiment 1.
\(N =\) 7 participants were excluded because they had less than 50\%
accurate responses on the unambiguous training trials. The final sample
consisted of \(N =\) 93 participants.

\subsubsection{Stimuli and Procedure}\label{stimuli-and-procedure}

We used the same visual stimuli as in Experiment 1. We also used the
same auditory stimuli, but we convolved each item with Brown noise of
amplitude 1 using the free sound editor Audacity (2.1.2). The average
signal-to-noise ratio was - 4.4 dB. The procedure was exactly the same
as in the previous experiment, except that the test stimuli (but not the
target) were presented with the new noisy auditory stimuli.

\subsection{Results}\label{results}

The analysis is similar to the analysis we did in Experiment 1.

\subsubsection{Unimodal condition}\label{unimodal-condition}

We fit a model for each modality. Figure \ref{fig:unimodal} shows human
responses together with their best fits. The visual data is a
replication of the visual data in Experiment 1. The auditory data, in
contrast, were flatter, showing more uncertainty.

\subsubsection{Bimodal condition}\label{bimodal-condition-2}

We used the values derived from the unimodal condition to construct the
visual, auditory and optimal models. In addition, we fit a descriptive
model which allowed us to assess real human performance in this
condition. Figure \ref{fig:bimodal} shows that, similar to Experiment 1,
the optimal model explained more variance than the auditory and visual
models did (note, however, that the visual model explained more variance
than the auditory model did). Also similar to Experiment 1, the values
of the variances increased in the bimodal condition (Table
\ref{tab:param}).

\subsubsection{Cue combination}\label{cue-combination-1}

Here we investigated whether the observed increase in the auditory and
visual variances affected the relative weighting of the corresponding
modalities. Figure~\ref{fig:bias} (top) shows that the participants'
decision threshold deviated from optimality and that this deviation was
biased towards the visual modality (the non-noisy modality). Indeed
non-parametric resampling of the data showed a decrease in the value of
the slope in the descriptive model compared to the optimal model
(Figure~\ref{fig:bias}, bottom).

\subsection{Discussion}\label{discussion-1}

Experiment 2 tested audio-visual combination in the case where the
auditory input was noisy. We found, similar to Experiment 1, that the
optimal model explained more variance than the auditory or the visual
models did. In other words, despite additional noise, participants still
used information from the noisy modality to recognize words. We also
found a similar discrepancy between the descriptive and optimal models
as response randomness increased along both the auditory and the visual
modalities. As for the relative weighting, and contrary to Experiment 1
where modalities were weighted optimally, we found in this experiment
that the visual modality had a greater weight than what was expected
from its relative reliability. This situation corresponds to the second
case of sub-optimality described in Figure \ref{fig:subOptim}.

Whereas in Experiment 2 we tested the case of added background noise to
the auditory modality, in Experiment 3 we test the case of added noise
to the visual modality.

\section{Experiment 3}\label{experiment-3}

Similar to Experiment 2, we were interested to know if participants
would treat additional uncertainty as predicted by the optimal model,
that is, according to the following weighting scheme:

\[\beta_a \propto \frac{1}{\sigma^2_{A}} = \frac{1}{\sigma^2_{A_C}+\sigma^2_{A_N}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}} = \frac{1}{\sigma^2_{V_C}+\sigma^2_{V_N} + \sigma^2_{V_E}}.\]

The alternative hypothesis is that noise in the visual modality would
lead to a preference for the auditory input, just like noise in the
auditory modality lead to a preference for the visual input in
Experiment 2.

\subsection{Methods}\label{methods-2}

\subsubsection{Participants}\label{participants-2}

A planned sample of \(N =\) 100 participants was recruited online
through Amazon Mechanical Turk. We used the same exclusion criteria as
in both previous experiments. \(N =\) 2 participants were excluded
because they reported having a technical problem, and \(N =\) 10
participants were excluded because they had less than 50\% accurate
responses on the unambiguous training trials. The final sample consisted
of \(N =\) 88 participants.

\subsubsection{Stimuli and Procedure}\label{stimuli-and-procedure-1}

We used the same auditory stimuli as in Experiment 1. We also used the
same visual stimuli, but we blurred the tokens using the free image
editor GIMP (2.8.20). We used a Gaussian blur with a
radius\footnote{A feature that modulates the intensity of the blur.} of
10 pixels. The experimental procedure was exactly the same as in the
previous Experiments.

\subsection{Results}\label{results-1}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_r2_files/figure-latex/bias-1} \caption{Modality preference is characterized as a deviation from the optimal decision threshold. A) The decision thresholds of both the optimal and the descriptive models (solid red and black lines, respectively). Deviation from optimality is compared to two hypothetical cases of modality preference. In these cases, deviation from optimality is due to over-lying on the visual or the auditory input by a factor of 2 (green and blue dotted lines, respectively). B) An alternative way to represent the same data. Each point represents the value of the decision threshold's slope derived from the descriptive model relative to that of the optimal model (log-scaled). The lines represent the optimal case as well as the two hypothetical cases of modality preference. Error bars represent 95\% confidence intervals over the distribution obtained through non-parametric resampling.}\label{fig:bias}
\end{figure}

\subsubsection{Unimodal conditions}\label{unimodal-conditions-2}

Figure \ref{fig:unimodal} shows responses in the unimodal conditions as
well as the corresponding fits. The auditory data is a replication of
the auditory data in Experiment 1. As for the visual data, we found
that, in contrast to Experiment 1 and 2, responses were flatter, showing
much more uncertainty.

\subsubsection{Bimodal condition}\label{bimodal-condition-3}

Figure \ref{fig:bimodal} shows that almost all the variance was captured
by the auditory model alone, the addition of visual information in the
optimal model did not improve the prediction of human responses. Similar
to Experiments 1 and 2, the values of the variances increased in the
bimodal condition (Table \ref{tab:param}).

\subsubsection{Cue Combination}\label{cue-combination-2}

Figure \ref{fig:bias} indicates that the decision threshold was biased
towards the auditory modality (the non-noisy modality). Indeed
non-parametric resampling of the data showed an increase in the value of
the slope in the descriptive model compared to the optimal model
(Figure~\ref{fig:bias}).

\subsection{Discussion}\label{discussion-2}

Experiment 3 tested audi-visual combination in the case where the visual
input was noisy. Whereas in previous experiments the optimal model
explained more variance than the auditory or the visual models did, here
the auditory model alone explained almost all the variance. In other
words, though participants were sensitive to variation in the noisy
visual input when presented in isolation (as shown in Figure
\ref{fig:unimodal}), they tended to ignore this information when the
visual input was presented simultaneously with the auditory input (i.e.,
in the bimodal condition). Instead, they relied almost exclusively on
the non-noisy auditory
modality.\footnote{The reason why we saw this (floor) effect when we added noise to the visual modality (Experiment 3), and not when we added noise to the auditory modality (Experiment 2), is the fact that our visual stimuli were originally perceived less categorically and with less certainty than the auditory stimuli (see Experiment 1 in Figure 4). This fact made it more likely for the visual categorization function to become flat and uninformative after a few drops in precision due to noise on the one hand, and to the additional randomness induced by the bimodal presentation on the other hand.}

This finding corresponds to the third case of sub-optimality described
in Figure~\ref{fig:subOptim}. Indeed, precision dropped for both
modalities in the bimodal condition compared to the unimodal condition.
But the drop was much greater for the visual modality, resulting in a
much lower weight assigned to it than what is expected from the optimal
model. Therefore, just like participants over-relied on the visual
modality when the auditory modality was noisy (Experiment 2), they also
over-relied on the auditory modality when the visual modality was noisy
(Experiment 3).

So far we have studied the problem of cue combination at the population
level --- the models were fit to the data aggregated across all
participants. However, it is important to investigate individual
variability, especially for cases when we reported optimal cue
combination (i.e., Experiment 1). In fact, optimality at the population
level can be spurious if it is obtained only on average while
individuals have sub-optimal strategies (e.g., over-relying on the
visual or the auditory modalities). In Experiment 4, below, we examine
how the average cue combination relates to individual strategies.

\section{Experiment 4}\label{experiment-4}

As we noted earlier, we did not have enough statistical power in
Experiment 1 to fit a different model for each participant. Thus, here
we used a higher power design, allowing us to collect the number of data
points necessary to model cue combination at the individual level.

\subsubsection{Participants}\label{participants-3}

We recruited a planned sample of \(N=\) 50 participants from Amazon
Mechanical Turk. Only participants with US IP addresses and a task
approval rate above 99\% were allowed to participate. Participants were
excluded if they reported having experienced a technical problem of any
sort during the online experiment (\(N=\) 0), or if they had less than
75\% accurate responses on the unambiguous training trials (\(N=\) 7).
The final sample consisted of \(N =\) 43 participants. All participants
provided informed consent before taking the
experiment.\footnote{The sample size, exclusion criteria and the main analyses were pre-registered at https://osf.io/h7mzp/.}

\subsubsection{Stimuli}\label{stimuli-1}

We used the same stimuli as in Experiment 1.

\subsubsection{Design and Procedure}\label{design-and-procedure-1}

The design and procedure were similar to Experiment 1. There were,
however, two differences: 1) We increased the number of responses
elicited per subject from 70 to 300, and 2) we randomized the order of
the three blocks (i.e., visual-only, auditory-only, and audio-visual)
\emph{within} subject: Each participant saw the 3 blocks exactly 6
times, covering all possible ordering combinations. Unlike the
between-subject randomization that we used in Experiment 1, this choice
allowed us to avoid a possible confound linked to the order of exposure.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_r2_files/figure-latex/weights-1} \caption{The histogram shows the distribution of the participants' predicted (i.e., optimal) cue weighting relative the observed (i.e., descriptive) weighting (see Figure 3 for the details). The density plot shows the distribution of simulated data sampled from the population-level probabilistic model. The dashed lines represent 95\% confidence interval on this simulated distribution. Optimal behavior is exactly obtained when the value of the relative cue weighting is 1 (red solid line). Participants whose values are outside the confidence interval of the simulated distribution can be understood to be over-relying on the visual modality (left side) or on the auditory modality (right side) beyond sampling-related variation (i.e., with p < 0.05).}\label{fig:weights}
\end{figure}

\subsection{Results}\label{results-2}

\subsubsection{Unimodal and Bimodal
conditions}\label{unimodal-and-bimodal-conditions}

In order to replicate the analysis of Experiment 1, we started by
fitting population-level models to the aggregated data. Indeed, we found
that the results --- as shown in Figure \ref{fig:unimodal}, Figure
\ref{fig:bimodal}, and Table \ref{tab:param} --- mirror closely the
patterns obtained in Experiment 1.

\subsubsection{Cue combination}\label{cue-combination-3}

We analyzed the cue combination strategies at the individual level. For
each participant, we computed the optimal weighting,
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\), relative to the observed (i.e.,
descriptive) weighting, \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). We show
the resulting distribution in Figure \ref{fig:weights}. We note first
that the distribution has a rather unimodal shape, centered around the
optimal cue combination strategy. This finding rules out the hypothesis
that optimality at the population level is a spurious finding, i.e.,
only obtained by aggregating over various sub-optimal strategies.

In addition, we asked whether the observed variance in the individual
distribution was due to the randomness inherent to the process of
sampling from a probabilistic model or whether it corresponded to a real
between-subject variability induced by different cue weighting
strategies. We simulated responses through sampling from the
population-level models and we computed the resulting distribution of
cue weighting for each simulated individual leading to the density
function in Figure \ref{fig:weights}. We can observe that most empirical
values fall within the 95\% confidence interval of the simulated
density, showing that this part of the variance can be due to mere
sampling variation. However, a few participants had values outside this
interval, indicating that they systematically over-relied on the visual
modality (\(N=\) 5, \(p < 0.05\)) or the auditory modality (\(N=\) 6,
\(p < 0.05\)).

\subsection{Discussion}\label{discussion-3}

This experiment was an extension to Experiment 1. Through collecting
denser data per subject, we were able to analyze the cue combination
optimality, not only at the population level but also at the individual
level. The population-level analysis replicated the results of
Experiment 1. The individual-level analysis showed that the distribution
of cue combination scores had a unimodally-shaped distribution centered
around the optimal combination, thus reflecting genuine cue combination
at the individual level. That said, the variance of this distribution
indicates that a few participants tended to over-rely on the auditory
modality and others tended to over-rely on the visual modality beyond
sampling-related randomness.

\section{General Discussion}\label{general-discussion}

In the current paper, we explored word recognition under uncertainty
about both words and their referents. We conducted an ideal observer
analysis of this task whereby a model provided predictions about how
information from each modality should be combined in an optimal fashion.
The predictions of the model were tested in a series of four experiments
where instances of both the form and the meaning were ambiguous with
respect to their category membership only (Experiment 1 and 4), when
instances of the form were perturbed with additional background noise
(Experiment 2), and when instances of the referent were perturbed with
additional visual noise (Experiment 3). We discuss the findings of these
studies first with respect to our ideal observer model and inferences
about optimality and second with respect to their implications for word
identification more generally.

\subsection{Patterns of optimality and
sub-optimality}\label{patterns-of-optimality-and-sub-optimality}

In all of our experiments, and when compared to the predictions of the
visual or the auditory models, participants generally relied on both
modalities to make their decisions in the bimodal condition. Indeed, in
Experiment 1 and 2, the optimal model accounted for more variance in
mean responses than the auditory or the visual models did. In Experiment
3, participants appeared to rely on one modality, but this was likely a
floor effect, due to the fact that noise made the visual input barely
perceptible. Further, in Experiment 1 and 4, which did not involve
background noise, participants not only relied on both modalities but
generally weighted these modalities according to the predictions of the
optimal model, that is, according to their relative reliability. At the
individual level, Experiment 4 showed that most participants were
near-optimal. Only a few subjects over-relied on the auditory or visual
modalities beyond sampling errors.

Despite this overall near-optimal behavior, we documented two major
cases of sub-optimality. First, in all experiments, the variance
associated with each modality increased in the bimodal condition
compared to the unimodal conditions: Participants responded slightly
more randomly. This increase in randomness could be due to limitations
on cognitive resources: Processing two separate --- and perceptually
uncorrelated --- cues instead of one cue (as in the unimodal case) is
likely to place extra demands on working memory, causing general
performance to drop (see Mattys \& Wiget, 2011).

Previous research has found similar cases of sub-optimal behavior. For
instance, studies that have explored the identification of ambiguous,
newly learned pairs of word-referent associations have reported what
appears to be a decrease in speech perception acuity in both children
(Stager \& Werker, 1997) and adults (Pajak, Creel, \& Levy, 2016). In
agreement with the findings of our study, Hofer and Levy (2017)
characterized the apparent reduction in perceptual acuity as an increase
in the noise variance of the auditory modality. Our findings, besides
providing more evidence to this documented fact, suggest that the
reduction in perceptual acuity may occur simultaneously in both the
auditory \emph{and} the visual modalities.

The second case of sub-optimality is related to how participants
weighted the cues from the visual and the auditory modalities in a noisy
context. In contrast to Experiment 1 and 4 where the cue combination was
indistinguishable from the optimal predictions, results of Experiment 2
and 3 suggested that participants had a systematic preference for the
non-noisy modality. This finding is reminiscent of the fact that humans
tend to compensate for a degraded speech signal by relying more on other
sources of information such as the accompanying visual cues, the
semantic/syntactic context, or the top-down expectations. This kind of
compensation has been observed with adults (Mattys et al., 2012;
Tanenhaus et al., 1995), and recent evidence suggests that it starts in
childhood (K. MacDonald, Marchman, Fernald, \& Frank, 2018; Yurovsky,
Case, \& Frank, 2017).

Generally speaking, previous experimental studies have not
differentiated between an optimal compensatory strategy (i.e., relying
more on the alternative source while using all information still
available in the distorted signal), and a sub-optimal strategy (i.e.,
relying more on the alternative source while ignoring at least some of
the information still available in the distorted signal), however. The
formal approach followed in this paper allowed us to tease apart these
two possibilities, and our analysis supports the sub-optimal
compensatory strategy: The preference for the non-noisy modality is
above and beyond what can be explained by the relative reliability
alone, meaning that the participants tend to ignore at least part of the
information still available in the noisy modality.

This second case of sub-optimal behavior may be related to the fact that
language understanding under degraded conditions is cognitively more
taxing than language understanding under normal conditions (Mattys et
al., 2012; Peelle, 2018; Rönnberg, Rudner, Lunner, Zekveld, \& others,
2010). Perhaps these demands lead to sub-optimal behavior (i.e.,
over-reliance on the less noisy cue) as participants seek to minimize
cognitive effort. One could also explain this phenomenon in terms of the
metacognitive experience about the fluency with which information is
processed. The perceived perceptual fluency (e.g., the ease with which a
stimulus' physical identity can be identified) can affect a wide variety
of human judgments (see Schwarz, 2004 for a review). In particular,
variables that improve fluency tend to increase liking/preference
(Reber, Winkielman, \& Schwarz, 1998). In our case, the subjective
experience of lower fluency in the noisy modality might cause people to
underestimate information that can be extracted from this modality,
especially when presented simultaneously with a higher fluency
alternative.

\subsection{Word recognition in the
wild}\label{word-recognition-in-the-wild}

An important question to ask is how the combination mechanism --- as
revealed in our controlled study --- scales up to real-life situations.
Note that in order to test audio-visual cue combination under
uncertainty, we had to use a case of double ambiguity, that is, a case
where both the word forms (\enquote{ada}--\enquote{aba}) and the
referents (cat--dog) were similar and, thus, confusable. To what extent
does such a case occur in real languages? Cross-linguistic corpus
analyses suggest that lexical encoding tends, surprisingly, towards
double ambiguity in many languages (Dautriche, Mahowald, Gibson, \&
Piantadosi, 2017; Monaghan, Shillcock, Christiansen, \& Kirby, 2014;
Tamariz, 2008). For instance, Dautriche et al. (2017) analyzed 100
languages and found that words that are similar phonologically tend to
be similar semantically as well. These studies suggest that the case of
double uncertainty, though perhaps not pervasive, could be a real issue
in language as it increases the probability of confusability for many
words. That said, the inferences discussed here might play a more
significant role in naturalistic language comprehension when ambiguity
in both the form and/or the referent is induced by an \emph{external}
noisy context --- e.g., a very noisy party or a far away referent ---
even when these forms and referents are not confusable in normal
situations.

Though we only studied adult performance in this paper, the problem of
word recognition under uncertainty is likely more pressing for children.
In fact, young children have greater difficulties learning the meanings
of novel similar-sounding words (e.g., \enquote{bin} vs. \enquote{din}),
even when these words are uttered very clearly (Creel, 2012; Merriman \&
Schuster, 1991; Stager \& Werker, 1997; Swingley, 2016; White \& Morgan,
2008). Such similar-sounding words can be shown to be differentiated by
infants in simplified experimental settings (e.g., Yoshida, Fennell,
Swingley, \& Werker, 2009). Nevertheless, Swingley (2007) suggested that
the ability to make this differentiation is likely not mature in early
childhood; children's representations are almost certainly noisier than
the adults' representations and may also be encoded with lower
confidence. Thus, children even more than adults might benefit from
additional disambiguating cues during new word-referent encoding and
recognition.

A multi-modal cue combination strategy might help children not only
recognize words, but also refine their underlying phonological and
semantic representations in the process. Previous research in early word
learning has --- whether implicitly or explicitly --- largely treated
the process of refining the word form and of refining the word meaning
as following a linear timeline. However, developmental data reveal that
children do not wait to have completed the acquisition of word forms
before they start learning their meanings (Bergelson \& Swingley, 2012;
Tincoff \& Jusczyk, 1999). Rather, both form and meaning representations
develop in a parallel fashion. A few studies have already suggested the
possibility of an interaction between sound and meaning in early
acquisition. For instance, Waxman and Markow (1995) showed that labeling
various objects with the same name helps infants form the underlying
semantic category (but see Sloutsky \& Napolitano, 2003). And in the
opposite direction, Yeung and Werker (2009) showed that pairing similar
sounds with different objects can help infants enhance their sensitivity
to subtle phonological contrasts in their native language. The present
study offers a stepping stone towards a formal framework where these
sorts of sound-meaning interactions in development can be unified and
further explored.

Note, however, that though our framework allows in principle for the
study of how sound-meaning interaction influences the underlying
categories, we did not explore this question in the current task. In
fact, we used familiar categories both at the seamantic level (cat and
dog) and at the phonological level (the phonemes /p/ and /b/). This
choice allowed us to test the mechanims of cue integration separately
from the question of category learning. Using novel categories would
create a confound as participant would additionally have to learn the
boundaries of the novel categories, making it difficult for us to tease
apart how much of the suboptimal behavior is due to the difficulty of
audio-visual integration, and how much it is due to variability in
estimating the boundary of the novel categories.

Though our task deals primarily with the mechanism of word recognition,
the conclusions of the current study partially applies to word learning
as well. While the typical task in the word learning literature involves
mapping a new label to a novel object (e.g., Landau, Smith, \& Jones,
1988; Markman, 1991), several cases of word learning in the wild involve
familiar object (as in our task). For example, adults learning a second
language do map new labels to objects that are already familiar to them.
Besides, children come to the task of word learning with a fairly
refined knowledge about the semantic categories (e.g., Quinn, Eimas, \&
Rosenkrantz, 1993). That is, in several cases, they do not associate new
labels with objects that they are seeing for the first time. Rather,
they map labels onto concepts that they have refined over the course of
earlier exposures (e.g., Bloom, 2000).

Finally, our task was an adaptation of a task introduced originally by
Sloutsky and Napolitano (2003). This original work is part of a rich
literature on modality-dominance in coross-modal processing (e.g.,
Colavita, 1974; Robinson \& Sloutsky, 2010). Overall, this litartaure
suggests that, when presented with au audio-visual stimuli, adults pay
more attention to the visual component whereas children may pay more
attention to the auditory component (Barnhart, Rivera, \& Robinson,
2018; Hirst, Cragg, \& Allen, 2018). Our work shows that adults do not
systematicallty prefere one modality over the other. Rather, they
modulate this preference as a functon of noise (Experiment 2 and 3).
That said, it is important to keep in mind that our task differs in its
design and aim from the tasks used in this literature: While the
literature on modality-dominance deals with low-level perceptual
encoding, our task targets specifically high-level linguistic processing
and categorizaton. It is possible that the underlying mechanisms in both
cases are related (or even similar). However, we did not explore this
possibility in the current study.

\subsection{Limitations}\label{limitations}

One salient limitation of our current work is that we used a restricted
and highly simplified stimulus set. For the auditory modality, we used
speech categories that varied along a single acoustic dimension. While
this dimension might be sufficient to recognize words in our specific
case, in general the speech signal is far more complex, varying along
several acoustic/phonetic dimensions. Additionally, these dimensions may
be highly variable due to various kinds of speaker and context
differences.

Concerning the visual dimension, simulating meaningful variability has
been a notoriously difficult problem. Following previous studies
(Freedman et al., 2001; Havy \& Waxman, 2016; Sloutsky \& Fisher, 2004),
we used a visual continuum along a one-dimensional morph. This
simplification was motivated by the need to construct a multimodal input
where the auditory and visual components are parametrized in a
symmetrical fashion, allowing us to compare graded effects of auditory
and visual information on categorical judgment. Though such a visual
variability is clearly artificial (one does not encounters in real life
an animal that is, e.g., 30 \% dog and 70 \% cat), we assume that the
induced uncertainty form these visual stimuli has a similar effect on
word recognition as the uncertainty induced by more naturalistic
semantic variability.

It is an open question whether people use the same strategy in
controlled laboratory conditions and more naturalistic settings where
they have to deal with various levels of variability. An answer to this
question is likely to involve a multifaceted research approach that goes
beyond controlled experimentation. We believe that one fruitful approach
is to test computational mechanisms with an input that more accurately
represents the full extent of multimodal variability in the learning
environment (Dupoux, 2018; Fourtassi, Schatz, Varadarajan, \& Dupoux,
2014; Harwath, Torralba, \& Glass, 2016; B. C. Roy, Frank, DeCamp, \&
Roy, 2015).

Finally, though we used the term \enquote{word recognition}, our work is
only indirectly related to the literature about how a rich lexicon is
accessed (e.g., McClelland \& Elman, 1986). We have used this term in a
more specific way, describing access to a simplified lexicon made of two
novel, ambiguous words. Such a simplified experimental context is not
new and has been crucial to our understanding of early word learning and
recognition (e.g., Stager \& Werker, 1997).

\section{Conclusions and future research
directions}\label{conclusions-and-future-research-directions}

Our work used an ideal observer model to study word recognition under
audio-visual uncertainty. This framework enabled us not only to test
optimality but also to examine systematically how and by how much people
deviate from optimality in their combination strategies. Thus, our work
is part of a growing effort to go beyond optimality tests --- which have
limited explanatory power --- and use models that also allow us to
identify and explain various patterns of sub-optimality in human
behavior (Rahnev \& Denison, 2018).

While we focused on the case of arbitrary associations in novel word
recognition, it is possible to extend the framework to other cases such
as that of \textit{iconicity}, i.e., when there is a resemblance between
the sound of a word and its referent. Previous work has suggested that
iconicity, among other things, helps with learning (and generalizing the
meaning of) new words (see Dingemanse, Blasi, Lupyan, Christiansen, \&
Monaghan, 2015 for a review). Using the research strategy in this paper,
we can, for example, test whether iconicity has such an advantage
because it mitigates the sub-optimal patterns observed with more
arbitrary pairings.

Finally, though the current framework only characterizes adult novel
word recognition, it provides a first step towards a model where
developmental questions can also be investigated. For instance, future
work should explore whether children, like adults, use probabilistic
cues from both the auditory and the visual input to recognize ambiguous
words, the extent to which they combine these cues in an optimal
fashion, and whether this cue combination helps them learn words and
refine their early phonological and semantic representations.

\vspace{1em}

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/WordRec}}}
\vspace{1em}

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation.

\section{Disclosure statement}\label{disclosure-statement}

None of the authors have any financial interest or a conflict of
interest regarding this work and this submission.

\section{Appendix 1: derivation of the posterior (Equation
1)}\label{appendix-1-derivation-of-the-posterior-equation-1}

For an ideal observer, the probability of choosing category 2 when
presented with an audio-visual instance \(w = (a, v)\) is the posterior
probability of this category:

\[p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}\]

Which reduces to:

\[p(W_2 | w)=\frac{1}{1+\frac{p(w|W_1)}{p(w|W_2)} \frac{p(W_1)}{p(W_2)}}\]
In order to further simplify the quantity \(\frac{p(w|W_1)}{p(w|W_2)}\),
we use our assumption that the cues are uncorrelated:
\[p(w | W) = p(a,v| W) = p(a| A)p(v| V)\] Using the \(\log\)
transformation, we get:

\[ \ln(\frac{p(w |W_1)}{p(w|W_2)})=\ln(\frac{p(a|W_1)}{p(a|W_2)})+\ln(\frac{p(v|W_1)}{p(v|W_2)}) \]
Under the assumption that the categories are normally distributed and
that, within each modality, the categories have equal variances, we get
(after simplification):

\[\ln(\frac{p(a|W_1)}{p(a|W_2)})=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\times a+ \frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}\]

and similarly:

\[\ln(\frac{p(v|W_1)}{p(v|W_2)})=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}\times v+ \frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

When putting all these terms together, we obtain this final expression
for the posterior:
\[p(W_2 | w)=\frac{1}{1+(1+b)\exp(\beta_0+\beta_aa+\beta_vv)}\]

where

\[1+b=\frac{p(W_1)}{p(W_2)}\]
\[\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

\[\beta_a=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\]
\[\beta_v=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}.\]

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-anderson90}{}
Anderson, J. R. (1990). \emph{The adaptive character of thought}.
Hillsdale, NJ: Erlbaum.

\hypertarget{ref-Bankieris17}{}
Bankieris, K. R., Bejjanki, V., \& Aslin, R. N. (2017). Sensory
cue-combination in the context of newly learned categories.
\emph{Scientific Reports}, \emph{7}(1), 10890.

\hypertarget{ref-barnhart2018}{}
Barnhart, W. R., Rivera, S., \& Robinson, C. W. (2018). Different
patterns of modality dominance across development. \emph{Acta
Psychologica}, \emph{182}.

\hypertarget{ref-bates88}{}
Bates, D., \& Watts, D. (1988). \emph{Nonlinear regression analysis and
its applications}. Wiley.

\hypertarget{ref-bejjanki2011}{}
Bejjanki, V., Clayards, M., Knill, D., \& Aslin, R. (2011). Cue
integration in categorical tasks: Insights from audio-visual speech
perception. \emph{PLoS ONE}, \emph{6}.

\hypertarget{ref-bergelson2012}{}
Bergelson, E., \& Swingley, D. (2012). At 6 to 9 months, human infants
know the meanings of many common nouns. \emph{Proceedings of the
National Academy of Sciences}, \emph{109}(9).

\hypertarget{ref-bloom2000}{}
Bloom, P. (2000). \emph{How children learn the meanings of words}.
Cambridge, MA. MIT Press.

\hypertarget{ref-Campbell2008}{}
Campbell, R. (2008). The processing of audio-visual speech: Empirical
and neural bases. \emph{Philosophical Transactions of the Royal Society
of London B: Biological Sciences}, \emph{363}(1493), 1001--1010.

\hypertarget{ref-chater06}{}
Chater, N., \& Manning, C. D. (2006). Probabilistic models of language
processing and acquisition. \emph{Trends in Cognitive Sciences},
\emph{10}, 335--344.

\hypertarget{ref-clayard08}{}
Clayards, M., Tanenhaus, M., Aslin, R., \& Jacobs, R. (2008). Perception
of speech reflects optimal use of probabilistic speech cues.
\emph{Cognition}, \emph{108}.

\hypertarget{ref-colavita1974}{}
Colavita, F. B. (1974). Human sensory dominance. \emph{Attention,
Perception, \& Psychophysics}, \emph{16}(2), 409--412.

\hypertarget{ref-Creel2012}{}
Creel, S. (2012). Phonological similarity and mutual exclusivity:
On-line recognition of atypical pronunciations in 3--5-year-olds.
\emph{Developmental Science}, \emph{15}(5), 697--713.

\hypertarget{ref-dautriche17}{}
Dautriche, I., Mahowald, K., Gibson, E., \& Piantadosi, S. (2017).
Wordform similarity increases with semantic similarity: An analysis of
100 languages. \emph{Cognitive Science}, \emph{41}(8), 2149--2169.

\hypertarget{ref-dingemanse2015}{}
Dingemanse, M., Blasi, D. E., Lupyan, G., Christiansen, M. H., \&
Monaghan, P. (2015). Arbitrariness, iconicity, and systematicity in
language. \emph{Trends in Cognitive Sciences}, \emph{19}(10), 603--615.

\hypertarget{ref-dupoux2018}{}
Dupoux, E. (2018). Cognitive science in the era of artificial
intelligence: A roadmap for reverse-engineering the infant
language-learner. \emph{Cognition}, \emph{173}, 43--59.

\hypertarget{ref-Eberhard1995}{}
Eberhard, K., Spivey-Knowlton, M. J., Sedivy, J. C., \& Tanenhaus, M.
(1995). Eye movements as a window into real-time spoken language
comprehension in natural contexts. \emph{Journal of Psycholinguistic
Research}, \emph{24}(6), 409--436.

\hypertarget{ref-edmiston2015}{}
Edmiston, P., \& Lupyan, G. (2015). What makes words special? Words as
unmotivated cues. \emph{Cognition}, \emph{143}, 93--100.

\hypertarget{ref-ernst02}{}
Ernst, M. O., \& Banks, M. S. (2002). Humans integrate visual and haptic
information in a statistically optimal fashion. \emph{Nature},
\emph{415}(6870), 429--433.

\hypertarget{ref-feldman2009}{}
Feldman, N., Griffiths, T., \& Morgan, J. (2009). The influence of
categories on perception: Explaining the perceptual magnet effect as
optimal statistical inference. \emph{Psychological Review},
\emph{116}(4), 752--782.

\hypertarget{ref-fourtassi2014b}{}
Fourtassi, A., Schatz, T., Varadarajan, B., \& Dupoux, E. (2014).
Exploring the relative role of bottom-up and top-down information in
phoneme learning. In \emph{Proceedings of the 52nd annual meeting of the
association for computational linguistics (volume 2: Short papers)}
(Vol. 2, pp. 1--6).

\hypertarget{ref-freedman2001}{}
Freedman, D., Riesenhuber, M., Poggio, T., \& Miller, E. and. (2001).
Categorical representation of visual stimuli in the primate prefrontal
cortex. \emph{Science}, \emph{291}.

\hypertarget{ref-Geisler2003}{}
Geisler, W. S. (2003). Ideal observer analysis. In \emph{The visual
neurosciences} (pp. 825--837). Cambridge, MA: MIT Press.

\hypertarget{ref-greenberg1957}{}
Greenberg, J. (1957). \emph{Essays in linguistics}. Chicago: University
of Chicago Press.

\hypertarget{ref-harwath2016}{}
Harwath, D., Torralba, A., \& Glass, J. (2016). Unsupervised learning of
spoken language with visual context. In \emph{Advances in neural
information processing systems} (pp. 1858--1866).

\hypertarget{ref-havy2016}{}
Havy, M., \& Waxman, S. (2016). Naming influences 9-month-olds'
identification of discrete categories along a perceptual continuum.
\emph{Cognition}, \emph{156}, 41--51.

\hypertarget{ref-hillenbrand1995}{}
Hillenbrand, J., Getty, L. A., Clark, M. J., \& Wheeler, K. (1995).
Acoustic characteristics of american english vowels. \emph{Journal of
the Acoustical Society of America}, \emph{97}.

\hypertarget{ref-hirst2018}{}
Hirst, R. J., Cragg, L., \& Allen, H. A. (2018). Vision dominates
audition in adults but not children: A meta-analysis of the colavita
effect. \emph{Neuroscience \& Biobehavioral Reviews}, \emph{94}.

\hypertarget{ref-hofer2017}{}
Hofer, M., \& Levy, R. (2017). Modeling Sources of Uncertainty in Spoken
Word Learning. In \emph{Proceedings of the 39th Annual Meeting of the
Cognitive Science Society}.

\hypertarget{ref-kleinschmidt2015}{}
Kleinschmidt, D. F., \& Jaeger, T. F. (2015). Robust speech perception:
Recognize the familiar, generalize to the similar, and adapt to the
novel. \emph{Psychological Review}, \emph{148}.

\hypertarget{ref-Knill04}{}
Knill, D., \& Pouget, A. (2004). The bayesian brain: The role of
uncertainty in neural coding and computation. \emph{Trends in
Neurosciences}, \emph{27}(12), 712--719.

\hypertarget{ref-kronrod2016}{}
Kronrod, Y., Coppess, E., \& Feldman, N. (2016). A unified account of
categorical effects in phonetic perception. \emph{Psychonomic Bulletin
\& Review}, \emph{23}(6).

\hypertarget{ref-kuhl1991}{}
Kuhl, P. K. (1991). Human adults and human infants show a ``perceptual
magnet effect'' for the prototypes of speech categories, monkeys do not.
\emph{Perception \& Psychophysics}, \emph{50}(2), 93--107.

\hypertarget{ref-landau1988}{}
Landau, B., Smith, L., \& Jones, S. (1988). The importance of shape in
early lexical learning. \emph{Cognitive Development}, \emph{3}.

\hypertarget{ref-lupyan2012}{}
Lupyan, G., \& Thompson-Schill, S. L. (2012). The evocative power of
words: Activation of concepts by verbal and nonverbal means.
\emph{Journal of Experimental Psychology: General}, \emph{141}(1), 170.

\hypertarget{ref-macdonald2018}{}
MacDonald, K., Marchman, V., Fernald, A., \& Frank, M. C. (2018). Adults
and preschoolers seek visual information to support language
comprehension in noisy environments. In \emph{Proceedings of the 40th
Annual Conference of the Cognitive Science Society}.

\hypertarget{ref-markman1991}{}
Markman, E. M. (1991). The whole-object, taxonomic, and mutual
exclusivity assumptions as initial constraints on word meanings. In S.
A. Gelman \& J. P. Byrnes (Eds.),.

\hypertarget{ref-marr1982}{}
Marr, D. (1982). \emph{Vision}. WH Freeman.

\hypertarget{ref-mattys11}{}
Mattys, S. L., \& Wiget, L. (2011). Effects of cognitive load on speech
recognition. \emph{Journal of Memory and Language}, \emph{65}(2),
145--160.

\hypertarget{ref-mattys12}{}
Mattys, S. L., Davis, M. H., Bradlow, A. R., \& Scott, S. K. (2012).
Speech recognition in adverse conditions: A review. \emph{Language and
Cognitive Processes}, \emph{27}(7-8), 953--978.

\hypertarget{ref-mcclelland1986}{}
McClelland, J. L., \& Elman, J. L. (1986). The trace model of speech
perception. \emph{Cognitive Psychology}, \emph{18}(1).

\hypertarget{ref-mcgurk1976}{}
McGurk, H., \& MacDonald, J. (1976). Hearing lips and seeing voices.
\emph{Nature}, \emph{264}, 746--748.

\hypertarget{ref-medina2011}{}
Medina, T., Snedeker, J., Trueswell, J., \& Gleitman, L. (2011). How
words can and cannot be learned by observation. \emph{Proceedings of the
National Academy of Sciences}, \emph{108}(22), 9014.

\hypertarget{ref-Merriman91}{}
Merriman, W., \& Schuster, J. (1991). Young children's disambiguation of
object name reference. \emph{Child Development}, \emph{62}(6),
1288--1301.

\hypertarget{ref-Monaghan2014}{}
Monaghan, P., Shillcock, R. C., Christiansen, M. H., \& Kirby, S.
(2014). How arbitrary is language? \emph{Philosophical Transactions of
the Royal Society of London B: Biological Sciences}, \emph{369}(1651).

\hypertarget{ref-Norris08}{}
Norris, D., \& McQueen, J. M. (2008). Shortlist B: A bayesian model of
continuous speech recognition. \emph{Psychological Review},
\emph{115}(2), 357--395.

\hypertarget{ref-pajak2016}{}
Pajak, B., Creel, S., \& Levy, R. (2016). Difficulty in learning
similar-sounding words: A developmental stage or a general property of
learning? \emph{Journal of Experimental Psychology: Learning, Memory,
and Cognition}, \emph{42}(9).

\hypertarget{ref-peelle2018}{}
Peelle, J. E. (2018). Listening effort: How the cognitive consequences
of acoustic challenge are reflected in brain and behavior. \emph{Ear and
Hearing}, \emph{39}(2), 204.

\hypertarget{ref-pinker1989}{}
Pinker, S. (1989). \emph{Learnability and cognition: The acquisition of
argument structure}. Cambridge, MA: MIT press.

\hypertarget{ref-quinn93}{}
Quinn, P. C., Eimas, P. D., \& Rosenkrantz, S. L. (1993). Evidence for
representations of perceptually similar natural categories by
3-month-old and 4-month-old infants. \emph{Perception}, \emph{22}(4),
463--475.

\hypertarget{ref-rahnev2018}{}
Rahnev, D., \& Denison, R. N. (2018). Suboptimality in perceptual
decision making. \emph{Behavioral and Brain Sciences}, 1--107.

\hypertarget{ref-reber98}{}
Reber, R., Winkielman, P., \& Schwarz, N. (1998). Effects of perceptual
fluency on affective judgments. \emph{Psychological Science},
\emph{9}(1), 45--48.

\hypertarget{ref-robinson2010}{}
Robinson, C. W., \& Sloutsky, V. (2010). Development of cross-modal
processing. \emph{Wiley Interdisciplinary Reviews: Cognitive Science},
\emph{1}.

\hypertarget{ref-roy2015}{}
Roy, B. C., Frank, M. C., DeCamp, M., P., \& Roy, D. (2015). Predicting
the birth of a spoken word. \emph{Proceedings of the National Academy of
Sciences}, \emph{112}.

\hypertarget{ref-ronnberg2010}{}
Rönnberg, J., Rudner, M., Lunner, T., Zekveld, A. A., \& others. (2010).
When cognition kicks in: Working memory and speech understanding in
noise. \emph{Noise and Health}, \emph{12}(49), 263.

\hypertarget{ref-saussure1916}{}
Saussure, F. (1916). \emph{Course in general linguistics.} New York:
McGraw-Hill.

\hypertarget{ref-schwarz2004}{}
Schwarz, N. (2004). Metacognitive experiences in consumer judgment and
decision making. \emph{Journal of Consumer Psychology}, \emph{14}(4),
332--348.

\hypertarget{ref-sloutsky2004}{}
Sloutsky, V., \& Fisher, A. V. (2004). Induction and categorization in
young children: A similarity-based model. \emph{Journal of Experimental
Psychology: General}, \emph{133}(2), 166.

\hypertarget{ref-sloutsky2003}{}
Sloutsky, V., \& Napolitano, A. (2003). Is a picture worth a thousand
words? Preference for auditory modality in young children. \emph{Child
Development}, \emph{74}.

\hypertarget{ref-smith08}{}
Smith, L. B., \& Yu, C. (2008). Infants rapidly learn word-referent
mappings via cross-situational statistics. \emph{Cognition},
\emph{106}(3), 1558--1568.

\hypertarget{ref-spivey2002}{}
Spivey, M. J., Tanenhaus, M., Eberhard, K., \& Sedivy, J. C. (2002). Eye
movements and spoken language comprehension: Effects of visual context
on syntactic ambiguity resolution. \emph{Cognitive Psychology},
\emph{45}(4), 447--481.

\hypertarget{ref-stager1997}{}
Stager, C. L., \& Werker, J. F. (1997). Infants listen for more phonetic
detail in speech perception than in word-learning tasks. \emph{Nature},
\emph{388}(6640).

\hypertarget{ref-suanda2014}{}
Suanda, S. H., Mugwanya, N., \& Namy, L. L. (2014). Cross-situational
statistical word learning in young children. \emph{Journal of
Experimental Child Psychology}, \emph{126}.

\hypertarget{ref-Swingley2007}{}
Swingley, D. (2007). Lexical exposure and word-form encoding in
1.5-year-olds. \emph{Developmental Psychology}, \emph{43}(2), 454--464.

\hypertarget{ref-Swingley2016}{}
Swingley, D. (2016). Two-year-olds interpret novel phonological
neighbors as familiar words. \emph{Developmental Psychology},
\emph{52}(7), 1011--1023.

\hypertarget{ref-Tamariz2008}{}
Tamariz, M. (2008). Exploring systematicity between phonological and
context-cooccurrence representations of the mental lexicon. \emph{The
Mental Lexicon}, \emph{3}(2).

\hypertarget{ref-Tanenhaus1995}{}
Tanenhaus, M., Spivey-Knowlton, M., Eberhard, K., \& Sedivy, J. (1995).
Integration of visual and linguistic information in spoken language
comprehension. \emph{Science}, \emph{268}(5217), 1632--1634.

\hypertarget{ref-tenenbaum11}{}
Tenenbaum, J., Kemp, C., Griffiths, T., \& Goodman, N. (2011). How to
grow a mind: Statistics, structure, and abstraction. \emph{Science},
\emph{331}(6022), 1279--1285.

\hypertarget{ref-tincoff1999}{}
Tincoff, R., \& Jusczyk, P. W. (1999). Some beginnings of word
comprehension in 6-month-olds. \emph{Psychological Science},
\emph{10}(2), 172--175.

\hypertarget{ref-vlach2013}{}
Vlach, H. A., \& Johnson, S. P. (2013). Memory constraints on infants'
cross-situational statistical learning. \emph{Cognition}, \emph{127}.

\hypertarget{ref-vouloumanos2008}{}
Vouloumanos, A. (2008). Fine-grained sensitivity to statistical
information in adult word learning. \emph{Cognition}, \emph{107}(2),
729--742.

\hypertarget{ref-vouloumanos2014}{}
Vouloumanos, A., \& Waxman, S. (2014). Listen up! Speech is for thinking
during infancy. \emph{Trends in Cognitive Sciences}, \emph{18}(12),
642--646.

\hypertarget{ref-vroomen2004}{}
Vroomen, J., Linden, S. van, Keetels, M., Gelder, B. de, \& Bertelson,
P. (2004). Selective adaptation and recalibration of auditory speech by
lipread information: Dissipation. \emph{Speech Communication},
\emph{44}.

\hypertarget{ref-waxman2009}{}
Waxman, S., \& Gelman, S. (2009). Early word-learning entails reference,
not merely associations. \emph{Trends in Cognitive Sciences},
\emph{13}(6).

\hypertarget{ref-waxman1995}{}
Waxman, S., \& Markow, D. (1995). Words as invitations to form
categories: Evidence from 12-to 13-month-old infants. \emph{Cognitive
Psychology}, \emph{29}(3), 257--302.

\hypertarget{ref-white2008b}{}
White, K., \& Morgan, J. (2008). Sub-segmental detail in early lexical
representations. \emph{Journal of Memory and Language}, \emph{59}.

\hypertarget{ref-yeung09}{}
Yeung, H., \& Werker, J. (2009). Learning words' sounds before learning
how words sound: 9-month-olds use distinct objects as cues to categorize
speech information. \emph{Cognition}, \emph{113}, 234--243.

\hypertarget{ref-yoshida2009}{}
Yoshida, K., Fennell, C., Swingley, D., \& Werker, J. (2009).
14-month-olds learn similar-sounding words. \emph{Developmental
Science}, \emph{12}.

\hypertarget{ref-yurovsky2015}{}
Yurovsky, D., \& Frank, M. C. (2015). An Integrative Account of
Constraints on Cross-Situational Learning. \emph{Cognition}, \emph{145}.

\hypertarget{ref-yurovsky2017}{}
Yurovsky, D., Case, S., \& Frank, M. C. (2017). Preschoolers flexibly
adapt to linguistic input in a noisy channel. \emph{Psychological
Science}, \emph{28}(1), 132--140.

\end{document}